---
title: "Gradient Analysis for soil factors and biota"
author: "Jacob Weverka"
date: "8/9/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vegan)
library(here)
library(janitor)
library(ggpubr)
```

```{r results="hide"}
knitr::knit(here("code/soil_data_cleanup.Rmd"))
knitr::knit(here("code/veg_data_cleanup.Rmd"))
knitr::knit(here("code/16S_cleanup.Rmd"))
```
##
NMS16S  = NMDS of 16S data
NMS_family = NMDS of bacterial phyla


```{r}

change[41:80,5:7] = change[1:40,5:7]

change_data = change %>% 
  mutate(id = case_when(depth_cm == "10 to 40" ~ paste(point_id, "-40", sep = ""),
                        depth_cm == "0 to 10" ~ paste(point_id, "-10", sep = ""))
         ) %>% 
  # filter(id %in% rownames(otu_standard),
         # id != "TOKA-099-10"
         # ) %>% 
  column_to_rownames("id")



change_data10 = change_data %>% 
  filter(depth_cm == "0 to 10") %>% 
  as.data.frame() %>% 
  filter(rownames(.) %in% rownames(new_com_table10)) %>% 
  filter(rownames(.) %in% rownames(otu_standard))

change_data40 = change_data %>% 
  filter(depth_cm == "10 to 40") %>% 
  as.data.frame() %>% 
  filter(rownames(.) %in% rownames(new_com_table40)) %>% 
  filter(rownames(.) %in% rownames(otu_standard))

all_change = bind_rows(change_data10, change_data40)



c_change = toka_c_points %>% 
  filter(rownames(.) %in% rownames(otu_standard)) %>% 
  filter(`2021` != 0)


otu_standard_match10 = otu_standard %>% 
  filter(rownames(.) %in% rownames(change_data10)) %>% 
  as.data.frame()

otu_standard_match40 = otu_standard %>% 
  filter(rownames(.) %in% rownames(change_data40)) %>% 
  as.data.frame()

otu_standard_match = bind_rows(otu_standard_match10, otu_standard_match40)

family_standard_match10 = family_standard %>%
  filter(rownames(.) %in% rownames(change_data10))

family_standard_match40 = family_standard %>%
  filter(rownames(.) %in% rownames(change_data40))

family_standard_match = bind_rows(family_standard_match10, family_standard_match40)

veg_data_match10 = new_com_table10 %>%
  filter(rownames(.) %in% rownames(change_data10))

veg_data_match40 = new_com_table40 %>%
  filter(rownames(.) %in% rownames(change_data40))
# 
# class_standard_match = class_standard %>% 
#   filter(rownames(.) %in% rownames(change_data))
# 
# otu_standard3 = otu_standard %>% 
#   filter(rownames(.) %in% rownames(c_change))

# family_standard3 = family_standard %>% 
#   filter(rownames(.) %in% rownames(c_change))
# 
# class_standard3 = class_standard %>% 
#   filter(rownames(.) %in% rownames(change_data))

#separate into surface and depth
change_mat10 = change %>% 
  filter(depth_cm == "0 to 10") %>% 
  column_to_rownames("point_id") %>% 
  filter(rownames(.) %in% canopy_sum$point_id) %>% 
  select(n_change, `2018`, n_2018, cn_change, change18, c_n_ratio, sand, silt, clay)

change_mat40 = change %>% 
  filter(depth_cm == "10 to 40") %>% 
  column_to_rownames("point_id") %>% 
  filter(rownames(.) %in% canopy_sum$point_id) %>% 
  select(n_change, `2018`, n_2018, cn_change, change18, c_n_ratio, sand, silt, clay)

#clean canopy data and select
canopy_mat10 = canopy_sum %>% 
  filter(year == 2018) %>% 
  mutate(point_id = paste(point_id, "-10", sep = "")) %>% 
  column_to_rownames("point_id") %>% 
  select(-year) %>% 
  mutate(DIAU = DIAU + DIAUA) %>% 
  select(-DIAUA)%>% 
  filter(rownames(.) %in% rownames(change_data10)) %>% 
  as.data.frame()

canopy_mat40 = canopy_sum %>% 
  filter(year == 2018) %>% 
  mutate(point_id = paste(point_id, "-40", sep = "")) %>% 
  column_to_rownames("point_id") %>% 
  select(-year) %>% 
  mutate(DIAU = DIAU + DIAUA) %>% 
  select(-DIAUA) %>% 
  filter(rownames(.) %in% rownames(change_data40)) %>% 
  as.data.frame()

z = bind_cols(change_data10, canopy_mat10, otu_standard_match10)

# c_10 = z[,c(6,10)]
c_10 = z %>% 
  select(`2018`, change18)
# rownames(c_10) = rownames(z)
# colnames(c_10) = "c_2018"
soil10 = z %>% 
  select(sand, silt, clay)

c_40 = change_data40 %>% 
  select(`2018`, change18)

soil40 = change_data40 %>% 
  select(sand, silt, clay)

soil_change =  change %>% 
  mutate(id = case_when(depth_cm == "10 to 40" ~ paste(point_id, "-40", sep = ""),
                        depth_cm == "0 to 10" ~ paste(point_id, "-10", sep = ""))
         ) %>% 
  column_to_rownames("id")


# change_data_vp10 = change_data10 %>% 

```

# Microbe analyses

## Diversity

```{r}
microbe_diversity = diversity(otu_standard_match) %>% 
  as.data.frame()

colnames(microbe_diversity) = "diversity"

microbe_richness = specnumber(otu_standard_match) %>% 
  as.data.frame()

colnames(microbe_richness) = "richness"

microbe_diversity_soil = microbe_diversity %>% 
  rownames_to_column() %>% 
  left_join(microbe_richness %>% rownames_to_column(), by = "rowname") %>% 
  left_join(soil_change %>% rownames_to_column(), by = "rowname") %>% 
  mutate(depth = case_when(depth_cm == "0 to 10" ~ "0-10 cm",
                           depth_cm == "10 to 40" ~ "10-40 cm"))

richness_sum = microbe_diversity_soil %>% 
  group_by(depth_cm) %>% 
  summarize(richness_mean = mean(richness), richness_sd = sd(richness))


microbe_diversity_soil10 = microbe_diversity_soil %>% 
  filter(depth_cm == "0 to 10")
microbe_diversity_soil40 = microbe_diversity_soil %>% 
  filter(depth_cm == "10 to 40")

```
###plot
```{r}


(asv_richness = ggplot(microbe_diversity_soil, aes(x = richness, y = `2018`)) +
  geom_point() +
  facet_wrap(. ~ depth) +
  geom_smooth(method = "lm", linetype = "dashed", color = "black") +
  # scale_color_manual(values = c("darkgreen", "purple"),  labels = c("0-10 cm", "10-40 cm"), name = "Depth") +
  theme_bw() +
  ylab("% SOC 2018") +
  xlab("ASV Richness 2018"))

ggsave(here::here("documents/figures/ASV_richness.png"), dpi = 300, plot = asv_richness, device = "png", width = 8, height = 6)

```

```{r}

(asv_richness_change = ggplot(microbe_diversity_soil, aes(x = richness, y = change21, color = depth_cm)) +
  geom_point() +
  facet_wrap(. ~ depth_cm) +
  geom_smooth(method = "lm") +
  scale_color_manual(values = c("darkgreen", "purple"), name = "Depth") +
  theme_bw() +
  ylab("Change % SOC 2018-2021") +
  xlab("ASV richness"))

ggsave(here::here("documents/figures/ASV_richness_change.png"), dpi = 300, plot = asv_richness_change, device = "png", width = 8, height = 6)


```

```{r}
divmod_surf = lm(`2018` ~ richness, data = microbe_diversity_soil)
summary(divmod_surf)

divmod_depth = lm(`2018` ~ richness, data = microbe_diversity_soil40)
summary(divmod_depth)

divmod_depth_change = lm(change18 ~ richness, data = microbe_diversity_soil40)
summary(divmod_depth_change)

divmod_surf_change = lm(change18 ~ richness, data = microbe_diversity_soil10)
summary(divmod_surf_change)


divmod_surf_future = lm(change21 ~ richness, data = microbe_diversity_soil10)
summary(divmod_surf_future)

divmod_depth_future = lm(change21 ~ richness, data = microbe_diversity_soil40)
summary(divmod_depth_future)

divmod_future = lm(change21 ~ richness + depth_cm, data = microbe_diversity_soil)
summary(divmod_future)

divmod_current = lm(`2018` ~ richness*depth_cm, data = microbe_diversity_soil)
summary(divmod_current)

divmod = lm(`2018` ~ richness + depth_cm, data = microbe_diversity_soil)
summary(divmod)

AIC(divmod, divmod_current)
```


## NMDS all depths

```{r}
NMS16S = metaMDS(otu_standard_match, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE, autotransform = FALSE)


```
```{r}


# dimcheckMDS(otu_standard_match, distance = "bray", k = 10, trymax = 20, autotransform = FALSE)
```


```{r}

scores_16s = as.data.frame(scores(NMS16S))
change_mat = change_data %>% 
  select(change18, `2018`, change21, depth_cm, cn_change, n_2018, c_n_ratio, sand, silt, clay) %>% 
  filter(rownames(.) %in% rownames(scores_16s))

change_mat_complete = change_mat %>% 
  select(depth_cm, change18, `2018`,  sand, silt, clay)



depth_scores = envfit(scores_16s, change_mat, nperm = 999, na.rm = T)
depth_scores_plot = as.data.frame(scores(depth_scores, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
                              rowname == "2018" ~ "% Organic C",
                              rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
                              rowname == "cn_change" ~ "C:N ratio change",
                              rowname == "n_2018" ~ "% Nitrogen",
                              rowname == "c_n_ratio" ~ "C:N ratio",
                              rowname == "change21" ~ "Future % Change",
                              rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname))

# = data.frame(depth_scores$vectors$arrows * sqrt(depth_scores$vectors$r))

depth_scores_complete = envfit(scores_16s, change_mat_complete, nperm = 999, na.rm = T)
depth_scores_plot_complete = as.data.frame(scores(depth_scores_complete, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
                              rowname == "2018" ~ "% Organic C",
                              rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
                              rowname == "cn_change" ~ "C:N ratio change",
                              rowname == "n_2018" ~ "% Nitrogen",
                              rowname == "c_n_ratio" ~ "C:N ratio",
                              rowname == "change21" ~ "Future % Change",
                              rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname))

fam_16_env = envfit(scores_16s, family_standard_match, nperm = 999, na.rm = TRUE)
fam_16_scores = as.data.frame(scores(fam_16_env, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(phylum = str_replace(rowname, "p__", ""))

scores_16s_data = as.data.frame(scores(NMS16S)) %>%
  rownames_to_column() %>% 
  left_join(change_mat_complete %>% rownames_to_column(), by = "rowname") %>% 
  mutate(depth_cm = as.factor(depth_cm))

surface_grp = scores_16s_data[scores_16s_data$depth_cm == "0 to 10",][chull(scores_16s_data[scores_16s_data$depth_cm == "0 to 10", c("NMDS1", "NMDS2")]),]

depth_grp = scores_16s_data[scores_16s_data$depth_cm == "10 to 40",][chull(scores_16s_data[scores_16s_data$depth_cm == "10 to 40", c("NMDS1", "NMDS2")]),]

hull_data = rbind(surface_grp, depth_grp)

NMDS.mean=aggregate(scores_16s_data[,1:2],list(group=scores_16s_data$depth_cm),mean)


veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
{
  theta <- (0:npoints) * 2 * pi/npoints
  Circle <- cbind(cos(theta), sin(theta))
  t(center + scale * t(Circle %*% chol(cov)))
}

df_ell <- data.frame()
for(g in levels(scores_16s_data$depth_cm)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(scores_16s_data[scores_16s_data$depth_cm==g,],
                                                   veganCovEllipse(cov.wt(cbind(NMDS1,NMDS2),wt=rep(1/length(NMDS1),length(NMDS1)))$cov,center=c(mean(NMDS1),mean(NMDS2)))))
                                ,group=g))
}

```


All points
```{r}
library(ggrepel)


ggplot(scores_16s_data, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = depth_cm), size = 4) +
  # geom_polygon(data = hull_data, aes(x = NMDS1, y = NMDS2, color = depth_cm, fill = depth_cm), alpha = 0.3) +
  geom_path(data = df_ell, aes(x = NMDS1, y = NMDS2, color = group), size = 1, linetype = 2) +
  scale_color_manual(values = c("darkgreen", "purple"), name = "Depth") + 
  scale_fill_manual(values = c("darkgreen", "purple"), name = "Depth") +
  # geom_label(aes(label = rowname)) +
  # geom_segment(data = depth_scores_plot, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               # color = "blue", 
               # arrow = arrow(length = unit(0.25, "cm"))) +
  # geom_text(data = depth_scores_plot, aes(x = NMDS1, y = NMDS2, label = variable),
            # size = 6) +
  geom_text(data = fam_16_scores, mapping = aes(x = NMDS1, y = NMDS2, label = phylum), color = "black", size = 3) +
  geom_text(aes(label = paste("stress = ", round(NMS16S$stress, 3)), x = 0.8, y = -.9), size = 4, color = "black") +
  theme_bw() +
  xlim(-.9, 1)

```

PERMANOVA for Depth
```{r}
bray_16S = vegdist(otu_standard_match, method = "bray")

adonis2(bray_16S ~ depth_cm, scores_16s_data)
```


```{r}
ggplot(scores_16s_data, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = depth_cm), size = 4) +
  scale_color_manual(values = c("darkgreen", "purple"), name = "Depth") +
  geom_segment(data = depth_scores_plot_complete, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "blue", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text(data = depth_scores_plot_complete, aes(x = NMDS1, y = NMDS2, label = variable),
            size = 6) +
  geom_text(data = fam_16_scores, mapping = aes(x = NMDS1, y = NMDS2, label = phylum), color = "red") +
  theme_bw()

```


```{r}
# plot(scores_16s)
# plot(depth_scores)
# plot(fam_16_scores, col = "red")
```

Separate by depth
```{r}
change_mat10 = change_mat %>% 
  filter(depth_cm == "0 to 10")

change_mat10_complete = change_mat_complete %>% 
  filter(depth_cm == "0 to 10")

change_mat40 = change_mat %>% 
  filter(depth_cm == "10 to 40")

change_mat40_complete = change_mat_complete %>% 
  filter(depth_cm == "10 to 40")
```

## NMDS 16S 10


```{r}


NMDS_16S_10 = metaMDS(otu_standard_match10, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE)
NMDS_16S_10

scores_16s_10 = as.data.frame(scores(NMDS_16S_10))
```


```{r}
scores_16s_10 = as.data.frame(scores(NMDS_16S_10))
ff = change_mat10_complete %>% filter(`2018` < 7)


# scores_16s_10 = scores_16s_10 %>% 
  # filter(rownames(.)  %in% rownames(ff))

# depth_scores10 = envfit(scores_16s_10, change_mat10, nperm = 999, na.rm = T)
# depth_scores_plot10 = as.data.frame(scores(depth_scores10, display = "vectors")) %>%
#   rownames_to_column() %>%
#    mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
#                               rowname == "2018" ~ "% Organic C (2018)",
#                               rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
#                               rowname == "cn_change" ~ "C:N ratio change",
#                               rowname == "n_2018" ~ "% Nitrogen",
#                               rowname == "c_n_ratio" ~ "C:N ratio",
#                               rowname == "change21" ~ "Future % Change",
# rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname))

depth_scores10_complete = envfit(scores_16s_10, change_mat10_complete %>% select(-sand, -silt, -clay), nperm = 999, na.rm = T)
depth_scores_plot10_complete = as.data.frame(scores(depth_scores10_complete, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
                              rowname == "2018" ~ "% Organic C (2018)",
                              rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
                              rowname == "cn_change" ~ "C:N ratio change",
                              rowname == "n_2018" ~ "% Nitrogen",
                              rowname == "c_n_ratio" ~ "C:N ratio",
                              rowname == "change21" ~ "Future % Change",
                              rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname)) %>% 
  filter(rowname %in% c( "change18", "2018" , "change21"))

# = data.frame(depth_scores$vectors$arrows * sqrt(depth_scores$vectors$r))

family_standard_10 = family_standard_match10 %>% 
  select_if(colSums(.) != 0)

fam_16_env10 = envfit(scores_16s_10, family_standard_10, nperm = 999)
fam_16_scores10 = as.data.frame(scores(fam_16_env10, display = "vectors")) %>% 
  rownames_to_column() %>% 
  left_join(as.data.frame(fam_16_env10$vectors$pvals) %>% rownames_to_column(), by = "rowname") %>% 
  # rownames_to_column() %>% 
  mutate(phylum = str_replace(rowname, "p__", "")) %>% 
  filter(`fam_16_env10$vectors$pvals` <= 0.05)

scores_16s_10 = as.data.frame(scores(NMDS_16S_10)) %>% 
  rownames_to_column() %>% 
  left_join(change_mat10_complete %>% rownames_to_column(), by = "rowname") 
# %>% 
#   filter(`2018` < 7)
```
###no outlier

```{r}
ff = change_mat10_complete %>% filter(`2018` < 7)

otu10_no_outlier = otu_standard_match10 %>% 
  filter(rownames(.) %in% rownames(ff))

family_no_outlier = family_standard_10 %>% 
  filter(rownames(.) %in% rownames(ff))

NMDS_16S_10_no = metaMDS(otu10_no_outlier, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE)
NMDS_16S_10_no

scores_16s_10_no = as.data.frame(scores(NMDS_16S_10_no))
```


```{r}
scores_16s_10_no = as.data.frame(scores(NMDS_16S_10_no))
ff = change_mat10_complete %>% filter(`2018` < 7)


# scores_16s_10 = scores_16s_10 %>% 
  # filter(rownames(.)  %in% rownames(ff))

# depth_scores10 = envfit(scores_16s_10, change_mat10, nperm = 999, na.rm = T)
# depth_scores_plot10 = as.data.frame(scores(depth_scores10, display = "vectors")) %>%
#   rownames_to_column() %>%
#    mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
#                               rowname == "2018" ~ "% Organic C (2018)",
#                               rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
#                               rowname == "cn_change" ~ "C:N ratio change",
#                               rowname == "n_2018" ~ "% Nitrogen",
#                               rowname == "c_n_ratio" ~ "C:N ratio",
#                               rowname == "change21" ~ "Future % Change",
# rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname))

depth_scores10_complete_no = envfit(scores_16s_10_no, ff %>% select(-sand, -silt, -clay), nperm = 999, na.rm = T)
depth_scores_plot10_complete_no = as.data.frame(scores(depth_scores10_complete_no, display = "vectors")) %>% 
  rownames_to_column() %>% 
   mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
                              rowname == "2018" ~ "% Organic C (2018)",
                              rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
                              rowname == "cn_change" ~ "C:N ratio change",
                              rowname == "n_2018" ~ "% Nitrogen",
                              rowname == "c_n_ratio" ~ "C:N ratio",
                              rowname == "change21" ~ "Future % Change",
                              rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname)) %>% 
  filter(rowname %in% c( "change18", "2018" , "change21"))

# = data.frame(depth_scores$vectors$arrows * sqrt(depth_scores$vectors$r))

family_standard_10_no = family_no_outlier%>% 
  select_if(colSums(.) != 0)

fam_16_env10_no = envfit(scores_16s_10_no, family_standard_10_no, nperm = 999)
fam_16_scores10_no = as.data.frame(scores(fam_16_env10_no, display = "vectors")) %>% 
  rownames_to_column() %>% 
  left_join(as.data.frame(fam_16_env10_no$vectors$pvals) %>% rownames_to_column(), by = "rowname") %>% 
  # rownames_to_column() %>% 
  mutate(phylum = str_replace(rowname, "p__", "")) %>% 
  filter(`fam_16_env10_no$vectors$pvals` <= 0.05)

scores_16s_10_no = as.data.frame(scores(NMDS_16S_10_no)) %>% 
  rownames_to_column() %>% 
  left_join(ff %>% rownames_to_column(), by = "rowname") 
# %>% 
#   filter(`2018` < 7)
```

### plot
```{r}
(plot16s_10 = ggplot(scores_16s_10, aes(NMDS1, NMDS2, color = `2018`)) +
  geom_point(size = 3) +
  geom_segment(data = depth_scores_plot10_complete, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "blue", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text(data = depth_scores_plot10_complete, aes(label = variable),
            size = 4, color = "black") +
  geom_text(data = fam_16_scores10, mapping = aes(x = NMDS1, y = NMDS2, label = phylum), color = "blue") +
  scale_color_viridis_b(name = "% Organic C") + 
  theme_bw())

```
```{r}
(plot16s_10_no = ggplot(scores_16s_10_no, aes(NMDS1, NMDS2, color = `2018`)) +
  geom_point(size = 3) +
  geom_segment(data = depth_scores_plot10_complete_no, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "blue", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text(data = depth_scores_plot10_complete_no, aes(label = variable),
            size = 4, color = "black") +
  geom_text(data = fam_16_scores10_no, mapping = aes(x = NMDS1, y = NMDS2, label = phylum), color = "blue") +
  scale_color_viridis_b(name = "% Organic C") + 
  theme_bw())
```

```{r}
depth_scores10_complete
depth_scores10_complete_no

# m = lm(`2018` ~ NMDS1 + NMDS2, data = scores_16s_10)
# summary(m)
# 
# 
# fam_16_env10
```

## Phylum analysis

```{r}

depth_key = all_change %>% 
  select(depth_cm) %>% 
  rownames_to_column()

fam_carbon_10 = family_standard_10 %>% 
  rownames_to_column() %>% 
  left_join(change_mat10_complete %>% select(`2018`, change18) %>% rownames_to_column(), by = "rowname") %>% 
  column_to_rownames("rowname")

family_standard_40 = family_standard_match40 %>% 
  select_if(colSums(.) != 0)

fam_carbon_40 = family_standard_40 %>% 
  rownames_to_column() %>% 
  left_join(change_mat40_complete %>% select(`2018`, change18) %>% rownames_to_column(), by = "rowname") %>% 
  column_to_rownames("rowname")

fam_carbon = family_standard %>% 
  rownames_to_column() %>% 
  left_join(change_mat_complete %>% select(`2018`, change18) %>% rownames_to_column(), by = "rowname") %>% 
  left_join(depth_key, by = "rowname") %>% 
  column_to_rownames("rowname") %>% 
  drop_na()




fam_pivot_10 = family_standard_match10 %>% 
  select_if(colSums(.) != 0) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname)

bottom_taxa = family_standard_match %>% 
  select(-p__Verrucomicrobia, -p__Acidobacteria, -p__Proteobacteria, -p__Bacteroidetes, -p__Planctomycetes, -p__Actinobacteria, -p__Gemmatimonadetes, -p__Firmicutes, -p__Chloroflexi) %>% 
  mutate(other = rowSums(.)) %>% 
  rownames_to_column() %>% 
  select(rowname, other)

top_taxa = family_standard_match %>% 
  select(p__Verrucomicrobia, p__Acidobacteria, p__Proteobacteria, p__Bacteroidetes, p__Planctomycetes, p__Actinobacteria, p__Gemmatimonadetes, p__Firmicutes, p__Chloroflexi) %>% 
  rownames_to_column() %>% 
  left_join(bottom_taxa, by = "rowname")

top_taxa_3 = family_standard_match %>% 
  select(p__Verrucomicrobia, p__Acidobacteria, p__Proteobacteria, p__Bacteroidetes,  p__Actinobacteria,  p__Gemmatimonadetes) %>% 
  mutate(total = rowSums(.))
  

top_taxa_pivot = top_taxa %>%
  left_join(depth_key, by = "rowname") %>% 
  mutate(point_id = substr(rowname, 1, nchar(rowname)-3)) %>% 
  select(-rowname) %>% 
  pivot_longer(c(-point_id, -depth_cm)) %>% 
  mutate(phylum = factor(str_replace(name, "p__", ""), levels = rev(c("Verrucomicrobia", "Acidobacteria", "Proteobacteria", "Bacteroidetes", "Planctomycetes", "Actinobacteria", "Gemmatimonadetes", "Firmicutes", "Chloroflexi", "other")) ))

# 
# levels = c(p__Verrucomicrobia, p__Acidobacteria, p__Proteobacteria, p__Bacteroidetes, p__Planctomycetes, p__Actinobacteria, p__Gemmatimonadetes, p__Firmicutes, p__Chloroflexi)

```


```{r}
library(calecopal)

top_taxa_pivot_10 = top_taxa_pivot %>% 
  filter(depth_cm == "0 to 10")

top_taxa_pivot_40 = top_taxa_pivot %>% 
  filter(depth_cm == "10 to 40")

ggplot(top_taxa_pivot, aes(x = point_id, y = value, fill = phylum)) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill="white")) +
  facet_grid(depth_cm ~ .) +
  scale_fill_manual(name = "Phylum", values = rev(cal_palette("superbloom3", n = 10, type = "continuous"))) +
  xlab("Sample Point") +
  ylab("Relative Abundance")


```
```{r}
depth_mod = lm(`2018` ~ depth_cm, data = fam_carbon)
depth_mod_log = lm(log(`2018`) ~ depth_cm, data = fam_carbon)
```


```{r}
verruco_model = lm((`2018`) ~ p__Verrucomicrobia*depth_cm, data = fam_carbon)
summary(verruco_model)


AIC(verruco_model, depth_mod)
```


```{r}
acid_model = lm((`2018`) ~ p__Acidobacteria*depth_cm, data = fam_carbon)
summary(acid_model)

AIC(acid_model, depth_mod)
```

```{r}
# proteo_model = lm(`2018` ~ p__Proteobacteria*depth_cm, data = (fam_carbon[!rownames(fam_carbon)%in% c("TOKA-089-40", "TOKA-089-10", "TOKA-064-40", "TOKA-064-10"),]))

proteo_model = lm((`2018`) ~ p__Proteobacteria*depth_cm, data = fam_carbon)
summary(proteo_model)

plot(proteo_model)

AIC(proteo_model, depth_mod)

ggplot(fam_carbon, aes(y = (`2018`), x = p__Proteobacteria, color = depth_cm)) +
  geom_point()
```
```{r}
bacteroidetes_model = lm(log(`2018`) ~ p__Bacteroidetes*depth_cm, data = fam_carbon)
summary(bacteroidetes_model)

AIC(bacteroidetes_model, depth_mod_log)
```
```{r}
plancto_mod = lm(`2018` ~ p__Planctomycetes*depth_cm, data = fam_carbon)
summary(plancto_mod)

AIC(plancto_mod, depth_mod)


```

## Microbe traits
```{r}

oligotroph = c("Verrucomicrobia", "Acidobacteria")
copiotroph = c("Proteobacteria", "Bacteroidetes", "Actinobacteria", "Gemmatimonadetes")

fam_carbon_function = fam_carbon %>% 
  mutate(oligotrophs = p__Verrucomicrobia + p__Acidobacteria,
         copiotrophs = p__Proteobacteria + p__Bacteroidetes + p__Actinobacteria + p__Gemmatimonadetes,
         c_o_ratio = copiotrophs/oligotrophs)

fam_carbon_function10 = fam_carbon_function %>% 
  filter(depth_cm == "0 to 10")

fam_carbon_function40 = fam_carbon_function %>% 
  filter(depth_cm == "10 to 40")


trait_mod10 = lm(`2018` ~ c_o_ratio, data = fam_carbon_function10)
summary(trait_mod10)

trait_mod40 = lm(`2018` ~ c_o_ratio, data = fam_carbon_function40)
summary(trait_mod40)

trait_mod = lm(log(`2018`) ~  c_o_ratio*depth_cm, data = fam_carbon_function)

trait_mod1 = lm(log(`2018`) ~  c_o_ratio*depth_cm, data = fam_carbon_function, weights = fam_carbon_function$copiotrophs + fam_carbon_function$oligotrophs)
summary(trait_mod)

trait_only = lm(log(`2018`) ~  c_o_ratio, data = fam_carbon_function)
summary(trait_only)

AIC(trait_mod, trait_mod1)



AIC(depth_mod_log, trait_mod, trait_only)

plot(trait_only)

trait_sum =  fam_carbon_function %>% 
  group_by(depth_cm) %>% 
  summarize(mean_c_o = mean(c_o_ratio),
            mean_C = mean(`2018`))

```
```{r}
funtion_surface = fam_carbon_function %>% 
  filter(depth_cm == "0 to 10")

funtion_depth = fam_carbon_function %>% 
  filter(depth_cm == "10 to 40")

sfm = lm(formula = log(`2018`) ~ c_o_ratio, data = funtion_surface)
summary(sfm)


dfm = lm(formula = log(`2018`) ~ c_o_ratio, data = funtion_depth)
summary(dfm)


trait_mod_noint = lm(log(`2018`) ~  c_o_ratio + depth_cm, data = fam_carbon_function)
summary(trait_mod_noint)

AIC(trait_mod, trait_mod_noint)


plot(trait_mod_noint)
```


```{r}
(function_16S_plot = ggplot(fam_carbon_function, (aes(x = c_o_ratio, y = log(`2018`)))) +
  geom_point(aes(color = depth_cm)) +
  theme_bw() +
  # facet_grid(. ~ depth_cm) +
  scale_color_manual(values = c("darkgreen", "purple"), name = "Depth") +
  theme(axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.background = element_rect(fill="white")) +
  # geom_smooth(data = fam_carbon_function %>% filter(depth_cm == "0 to 10"), method = "lm", fill = "darkgreen", alpha = 0.1) +
  geom_smooth(method = "lm", color = "black") +
  ylab("log(% Organic Carbon)") +
  xlab("Copiotroph:Oligotroph Ratio"))

ggsave(here::here("documents/figures/function_16S_plot.png"), dpi = 300, plot = function_16S_plot, device = "png", width = 8, height = 6)

```


## NMDS 16S 40

```{r}
ff40 = change_mat40_complete %>% filter(`2018` < 5)

otu40_no_outlier = otu_standard_match40 %>% 
  filter(rownames(.) %in% rownames(ff40))

family_no_outlier40 = family_standard_40 %>% 
  filter(rownames(.) %in% rownames(ff40))

NMDS_16S_40 = metaMDS(otu_standard_match40, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE)
NMDS_16S_40

# scores_16s_10 = as.data.frame(scores(NMDS_16S_10))

scores_16s_40 = as.data.frame(scores(NMDS_16S_40))
```


```{r}
scores_16s_40 = as.data.frame(scores(NMDS_16S_40))
# depth_scores40 = envfit(scores_16s_40, change_mat40, nperm = 999, na.rm = T)
# depth_scores_plot40 = as.data.frame(scores(depth_scores40, display = "vectors")) %>% 
#   rownames_to_column() %>% 
#   mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
#                               rowname == "2018" ~ "% Organic C",
#                               rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
#                               rowname == "cn_change" ~ "C:N ratio change",
#                               rowname == "n_2018" ~ "% Nitrogen",
#                               rowname == "c_n_ratio" ~ "C:N ratio",
#                               rowname == "change21" ~ "Future % Change",
#                               rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname))

depth_scores40_complete = envfit(scores_16s_40, change_mat40_complete %>% select(-sand, -silt, -clay), nperm = 999, na.rm = T)
depth_scores_plot40_complete = as.data.frame(scores(depth_scores40_complete, display = "vectors")) %>% 
  rownames_to_column() %>% 
   mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
                              rowname == "2018" ~ "% Organic C",
                              rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
                              rowname == "cn_change" ~ "C:N ratio change",
                              rowname == "n_2018" ~ "% Nitrogen",
                              rowname == "c_n_ratio" ~ "C:N ratio",
                              rowname == "change21" ~ "Future % Change",
                              rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname)) %>% 
  filter(rowname %in% c( "change18", "2018" , "change21"))
# = data.frame(depth_scores$vectors$arrows * sqrt(depth_scores$vectors$r))

family_standard_40 = family_standard_match40 %>% 
  select_if(colSums(.) != 0)

fam_16_env40 = envfit(scores_16s_40, family_standard_40, nperm = 999)
fam_16_scores40 = as.data.frame(scores(fam_16_env40, display = "vectors")) %>% 
  rownames_to_column() %>% 
  left_join(as.data.frame(fam_16_env40$vectors$pvals) %>% rownames_to_column(), by = "rowname") %>% 
  mutate(phylum = str_replace(rowname, "p__", "")) %>% 
  filter(`fam_16_env40$vectors$pvals` <= 0.05)
# 
# fam_16_scores10 = as.data.frame(scores(fam_16_env10, display = "vectors")) %>% 
#   rownames_to_column() %>% 
#   left_join(as.data.frame(fam_16_env10$vectors$pvals) %>% rownames_to_column(), by = "rowname") %>% 
#   # rownames_to_column() %>% 
#   mutate(phylum = str_replace(rowname, "p__", "")) %>% 
#   filter(`fam_16_env10$vectors$pvals` <= 0.05)


scores_16s_40 = as.data.frame(scores(NMDS_16S_40)) %>% 
  rownames_to_column() %>%
  left_join(change_mat40_complete %>% rownames_to_column(), by = "rowname")

```


### no outlier

```{r}
NMDS_16S_40_no = metaMDS(otu40_no_outlier, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE)
NMDS_16S_40_no
```


```{r}
scores_16s_40_no = as.data.frame(scores(NMDS_16S_40_no))


# scores_16s_40 = scores_16s_40 %>% 
  # filter(rownames(.)  %in% rownames(ff40))

# depth_scores40 = envfit(scores_16s_40, change_mat40, nperm = 999, na.rm = T)
# depth_scores_plot40 = as.data.frame(scores(depth_scores40, display = "vectors")) %>%
#   rownames_to_column() %>%
#    mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
#                               rowname == "2018" ~ "% Organic C (2018)",
#                               rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
#                               rowname == "cn_change" ~ "C:N ratio change",
#                               rowname == "n_2018" ~ "% Nitrogen",
#                               rowname == "c_n_ratio" ~ "C:N ratio",
#                               rowname == "change21" ~ "Future % Change",
# rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname))

depth_scores40_complete_no = envfit(scores_16s_40_no, ff40 %>% select(-sand, -silt, -clay), nperm = 999, na.rm = T)
depth_scores_plot40_complete_no = as.data.frame(scores(depth_scores40_complete_no, display = "vectors")) %>% 
  rownames_to_column() %>% 
   mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
                              rowname == "2018" ~ "% Organic C (2018)",
                              rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
                              rowname == "cn_change" ~ "C:N ratio change",
                              rowname == "n_2018" ~ "% Nitrogen",
                              rowname == "c_n_ratio" ~ "C:N ratio",
                              rowname == "change21" ~ "Future % Change",
                              rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname)) %>% 
  filter(rowname %in% c( "change18", "2018" , "change21"))

# = data.frame(depth_scores$vectors$arrows * sqrt(depth_scores$vectors$r))

family_standard_40_no = family_no_outlier40%>% 
  select_if(colSums(.) != 0)

fam_16_env40_no = envfit(scores_16s_40_no, family_standard_40_no, nperm = 999)
fam_16_scores40_no = as.data.frame(scores(fam_16_env40_no, display = "vectors")) %>% 
  rownames_to_column() %>% 
  left_join(as.data.frame(fam_16_env40_no$vectors$pvals) %>% rownames_to_column(), by = "rowname") %>% 
  # rownames_to_column() %>% 
  mutate(phylum = str_replace(rowname, "p__", "")) %>% 
  filter(`fam_16_env40_no$vectors$pvals` <= 0.05)

scores_16s_40_no = as.data.frame(scores(NMDS_16S_40_no)) %>% 
  rownames_to_column() %>% 
  left_join(ff40 %>% rownames_to_column(), by = "rowname") 
```

```{r}
# ggplot(scores_16s_40, aes(MDS1, MDS2)) +
#   geom_point() +
#   geom_segment(data = depth_scores_plot40, aes(x = 0, xend = MDS1, y = 0, yend = MDS2), 
#                color = "blue", 
#                arrow = arrow(length = unit(0.25, "cm"))) +
#   geom_text(data = depth_scores_plot40, aes(x = MDS1, y = MDS2, label = variable),
#             size = 6) +
#   geom_text(data = fam_16_scores40, mapping = aes(x = MDS1, y = MDS2, label = phylum), color = "red") +
#   theme_bw()

```

```{r}
depth_scores40_complete
depth_scores40_complete_no
```

###plot
```{r}
(plot16s_40 = ggplot(scores_16s_40, aes(NMDS1, NMDS2, color = `2018`)) +
  geom_point(size = 3) +
  geom_segment(data = depth_scores_plot40_complete, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "blue", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text(data = depth_scores_plot40_complete, aes(label = variable),
            size = 4, color = "black") +
  geom_text(data = fam_16_scores40, mapping = aes(x = NMDS1, y = NMDS2, label = phylum), color = "blue") +
  scale_color_viridis_c(name = "% Organic C") + 
  theme_bw())

```

```{r}

(plot16s_40_no = ggplot(scores_16s_40_no, aes(NMDS1, NMDS2, color = `2018`)) +
  geom_point(size = 3) +
  geom_segment(data = depth_scores_plot40_complete_no, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "blue", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text(data = depth_scores_plot40_complete_no, aes(label = variable),
            size = 4, color = "black") +
  geom_text(data = fam_16_scores40_no, mapping = aes(x = NMDS1, y = NMDS2, label = phylum), color = "blue") +
  scale_color_viridis_b(name = "% Organic C") + 
  theme_bw())



```
ggsave(here::here("documents/figures/veg_NMDS_all.png"), dpi = 300, plot = veg_NMDS_all, device = "png", width = 8, height = 12)

```{r}
(bact_NMDS_all = ggarrange(plot16s_10, plot16s_40, common.legend = F, legend = "right", ncol = 1)  +
  theme(axis.title = element_text(size = 12)))

(bact_NMDS_all_no_outlier = ggarrange(plot16s_10_no, plot16s_40_no, common.legend =F, legend = "right", ncol = 1)  +
  theme(axis.title = element_text(size = 12)))


ggsave(here::here("documents/figures/bact_NMDS_all.png"), dpi = 300, plot = bact_NMDS_all, device = "png", width = 8, height = 12)

ggsave(here::here("documents/figures/bact_NMDS_all_no_outlier.png"), dpi = 300, plot = bact_NMDS_all_no_outlier, device = "png", width = 8, height = 12)

```




```{r}
ggplot(scores_16s_40, aes(x = NMDS2, y = `2018`)) +
  geom_point()
```
```{r}
depth_scores40_complete

m = lm(`2018` ~ NMDS1 + NMDS2, data = scores_16s_40)
summary(m)

```
## NMDS Future All

```{r}


toka_c_otu_points = toka_c_points %>% 
  filter(!is.na(`2021`))

toka_future_covariates = toka_c_otu_points %>% 
  select(`2015`, `2018`, change18, change21, depth) %>% 
  rownames_to_column() %>% 
  left_join(change_mat %>% rownames_to_column(), by = c("rowname", "2018", "change18", "change21")) %>% 
  column_to_rownames("rowname")

otu_future = otu_standard_match %>% 
  filter(rownames(.) %in% rownames(toka_future_covariates))









```



```{r}

change_mat_future = toka_future_covariates %>% 
select(depth_cm, change18, change21, `2018`, sand, silt, clay) %>% 
  filter(rownames(.) %in% rownames(scores_16s))

  
```


```{r}

NMDS_16S_future = metaMDS(otu_future, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE)

NMDS_16S_future

```



```{r}


scores_16s_future = as.data.frame(scores(NMDS_16S_future))


depth_scores_future = envfit(scores_16s_future, change_mat_future, nperm = 999, na.rm = T)
depth_scores_plot_future = as.data.frame(scores(depth_scores_future, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
                              rowname == "2018" ~ "% Organic C",
                              rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
                              rowname == "cn_change" ~ "C:N ratio change",
                              rowname == "n_2018" ~ "% Nitrogen",
                              rowname == "c_n_ratio" ~ "C:N ratio",
                              rowname == "change21" ~ "Future % Change",
                              rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname))

# = data.frame(depth_scores$vectors$arrows * sqrt(depth_scores$vectors$r))



scores_16s_future = as.data.frame(scores(NMDS_16S_future)) %>% 
  rownames_to_column() %>% 
  left_join(change_mat_future %>% rownames_to_column(), by = "rowname")

```


All points
```{r}
library(ggrepel)

ggplot(scores_16s_future, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = depth_cm), size = 4) +
  scale_color_manual(values = c("darkgreen", "purple"), name = "Depth") +
  geom_segment(data = depth_scores_plot_future, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "blue", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text(data = depth_scores_plot_future, aes(x = NMDS1, y = NMDS2, label = variable),
            size = 6) +
  geom_text(data = fam_16_scores, mapping = aes(x = NMDS1, y = NMDS2, label = phylum), color = "red") +
  theme_bw()

```






```{r}
depth_scores_future

```



## NMDS Future 10

```{r}



otu_future10 = otu_future %>% 
  filter(rownames(.) %in% rownames(otu_standard_match10))


```



```{r}

change_mat_future10 = toka_future_covariates %>% 
select(depth_cm, change18, change21, `2018`, sand, silt, clay) %>% 
  filter(rownames(.) %in% rownames(otu_future10))

  
```


```{r}

NMDS_16S_future10 = metaMDS(otu_future10, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE)

NMDS_16S_future10

```



```{r}


scores_16s_future10 = as.data.frame(scores(NMDS_16S_future10))


depth_scores_future10 = envfit(scores_16s_future10, change_mat_future10 %>% select(-sand, -silt, -clay, -change18, -`2018`), nperm = 999, na.rm = T)
depth_scores_plot_future10 = as.data.frame(scores(depth_scores_future10, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
                              rowname == "2018" ~ "% Organic C",
                              rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
                              rowname == "cn_change" ~ "C:N ratio change",
                              rowname == "n_2018" ~ "% Nitrogen",
                              rowname == "c_n_ratio" ~ "C:N ratio",
                              rowname == "change21" ~ "Future % Change",
                              rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname))

# = data.frame(depth_scores$vectors$arrows * sqrt(depth_scores$vectors$r))



scores_16s_future10 = as.data.frame(scores(NMDS_16S_future10)) %>% 
  rownames_to_column() %>% 
  left_join(change_mat_future10 %>% rownames_to_column(), by = "rowname")

```


All points
```{r}
library(ggrepel)

ggplot(scores_16s_future10, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = change21), size = 4) +
  # scale_color_manual(values = c("darkgreen", "purple"), name = "Depth") +
  geom_segment(data = depth_scores_plot_future10, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "blue", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text(data = depth_scores_plot_future10, aes(x = NMDS1, y = NMDS2, label = variable),
            size = 6) +
  # geom_text(data = fam_16_scores, mapping = aes(x = NMDS1, y = NMDS2, label = phylum), color = "red") +
  theme_bw()

```
```{r}
depth_scores_future10
```


## NMDS Future 40



```{r}



otu_future40 = otu_future %>% 
  filter(rownames(.) %in% rownames(otu_standard_match40))


```



```{r}

change_mat_future40 = toka_future_covariates %>% 
select(depth_cm, change18, change21, `2018`, sand, silt, clay) %>% 
  filter(rownames(.) %in% rownames(otu_future40))

  
```


```{r}

NMDS_16S_future40 = metaMDS(otu_future40, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE)

NMDS_16S_future40

```



```{r}


scores_16s_future40 = as.data.frame(scores(NMDS_16S_future40))


depth_scores_future40 = envfit(scores_16s_future40, change_mat_future40 %>% select(-sand, -silt, -clay, -change18, -`2018`), nperm = 999, na.rm = T)
depth_scores_plot_future40 = as.data.frame(scores(depth_scores_future40, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(variable = case_when(rowname == "change18" ~ "% Carbon Change",
                              rowname == "2018" ~ "% Organic C",
                              rowname == "c_pc_change" ~ "% Carbon Change (Relative)",
                              rowname == "cn_change" ~ "C:N ratio change",
                              rowname == "n_2018" ~ "% Nitrogen",
                              rowname == "c_n_ratio" ~ "C:N ratio",
                              rowname == "change21" ~ "Future % Change",
                              rowname == "sand" | rowname == "silt" | rowname == "clay" ~ rowname))

# = data.frame(depth_scores$vectors$arrows * sqrt(depth_scores$vectors$r))



scores_16s_future40 = as.data.frame(scores(NMDS_16S_future40)) %>% 
  rownames_to_column() %>% 
  left_join(change_mat_future40 %>% rownames_to_column(), by = "rowname")

```


All points
```{r}
library(ggrepel)

ggplot(scores_16s_future40, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = change21), size = 4) +
  # scale_color_manual(values = c("darkgreen", "purple"), name = "Depth") +
  geom_segment(data = depth_scores_plot_future40, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "blue", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text(data = depth_scores_plot_future40, aes(x = NMDS1, y = NMDS2, label = variable),
            size = 6) +
  # geom_text(data = fam_16_scores, mapping = aes(x = NMDS1, y = NMDS2, label = phylum), color = "red") +
  theme_bw()

```

```{r}
depth_scores_future40
```










<!-- ```{r} -->


<!-- canopy_mat = canopy_sum %>%  -->
<!--   column_to_rownames("point_id") %>%  -->
<!--   select(-year) %>%  -->
<!--   mutate(DIAU = DIAU + DIAUA) %>%  -->
<!--   select(-DIAUA)%>%  -->
<!--   filter(rownames(.) %in% rownames(change_mat10)) -->


<!-- nms_canopy = metaMDS(canopy_mat, distance = "bray", k = 3, maxit = 999, trymax = 500) -->

<!-- scores_canopy = nms_canopy$species -->
<!-- scores_canopy_points = nms_canopy$points -->

<!-- canopy_env_scores = envfit(scores_canopy_points, change_mat10, nperm = 999, na.rm = TRUE) -->
<!-- canopy_env_scores -->

<!-- canopy_env_scores40 = envfit(scores_canopy_points, change_mat40, nperm = 999, na.rm = TRUE) -->


<!-- ``` -->

<!-- ```{r} -->
<!-- plot(nms_canopy) -->
<!-- plot(canopy_env_scores) -->
<!-- text(scores_canopy, labels = rownames(scores_canopy)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- canopy_cca = cca(class_standard3 ~ change18 + change21, data = change_mat10) -->

<!-- ``` -->



```{r}
# c_change_canopy = c_change %>% 
#   left_join(canopy_cover, by = c("point_name" = "point_id"))
# 
# m = lm(percent_cover ~ change18 + change21, data = c_change_canopy)
# summary(m)
# 
# m1 = lm(percent_cover ~ `2015` + `2018` + `2021`, data = c_change_canopy)
# summary(m1)
# 
#   cor(c_change_canopy$change21, c_change_canopy$`2021`)

```



# Veg analyses


```{r}
soil_change =  change %>% 
  mutate(id = case_when(depth_cm == "10 to 40" ~ paste(point_id, "-40", sep = ""),
                        depth_cm == "0 to 10" ~ paste(point_id, "-10", sep = ""))
         ) %>% 
  column_to_rownames("id")

com_table_fun_standard10 = com_table_fun %>% 
  filter(year == "2018") %>% 
  mutate(id = paste(point_id, 10, sep = "-")) %>% 
  column_to_rownames("id") %>% 
  select(-point_id, -year) %>% 
  decostand(method = "total") %>% 
  filter(rownames(.) %in% rownames(soil_change))

com_table_fun_standard40 = com_table_fun %>% 
  filter(year == "2018") %>% 
  mutate(id = paste(point_id, 40, sep = "-")) %>% 
  column_to_rownames("id") %>% 
  select(-point_id, -year) %>% 
  decostand(method = "total") %>% 
  filter(rownames(.) %in% rownames((soil_change)))

com_table10 = new_com_table10 %>% 
  filter(rownames(.) %in% rownames((soil_change)))

com_table40 = new_com_table40 %>% 
  filter(rownames(.) %in% rownames((soil_change)))

soil_veg_match10 = soil_change %>% 
  filter(rownames(.) %in% rownames(com_table_fun_standard10))  %>% 
  select(change18, `2018`, sand, silt, clay)

soil_veg_match40 = soil_change %>% 
  filter(rownames(.) %in% rownames(com_table_fun_standard40)) %>% 
  select(change18, `2018`, sand, silt, clay)

soil_veg_match10_future = soil_change %>% 
  filter(rownames(.) %in% rownames(com_table_fun_standard10))  %>% 
  select(change18, change21, `2018`, sand, silt, clay) %>% 
  filter(!is.na(change21))

soil_veg_match40_future = soil_change %>% 
  filter(rownames(.) %in% rownames(com_table_fun_standard40)) %>% 
  select(change18, change21, `2018`, sand, silt, clay) %>% 
  filter(!is.na(change21))

soil10 = soil_veg_match10 %>% 
  rownames_to_column() %>% 
  mutate(point_id = str_sub(rowname, 1, 8)) %>% 
  select(point_id, change18, `2018`)

soil40 = soil_veg_match40 %>% 
  rownames_to_column() %>% 
  mutate(point_id = str_sub(rowname, 1, 8)) %>% 
  select(point_id, change18, `2018`)

soil_veg = soil10 %>% 
  left_join(soil40, by = "point_id", suffix = c("_10", "_40")) %>% 
  column_to_rownames("point_id")

com_table = com_table10 %>% 
  rownames_to_column() %>% 
  mutate(point_id = str_sub(rowname, 1, 8)) %>% 
  select(-rowname) %>% 
  column_to_rownames("point_id")
  
com_table_fun_all = com_table_fun_standard10 %>% 
  rownames_to_column() %>% 
  mutate(point_id = str_sub(rowname, 1, 8)) %>% 
  select(-rowname) %>% 
  column_to_rownames("point_id")

com_table_future = com_table_10_future %>% 
  rownames_to_column() %>% 
  mutate(point_id = str_sub(rowname, 1, 8)) %>% 
  select(-rowname) %>% 
  column_to_rownames("point_id")

soil_veg_future = soil_change %>% 
  filter(rownames(.) %in% rownames(com_table_10_future) | rownames(.) %in% rownames(com_table_40_future)) %>% 
  select(point_id, depth_cm, change21) %>% 
  pivot_wider(values_from = change21, names_from = depth_cm) %>% 
  rename(surface = `0 to 10`,
         depth = `10 to 40`)

com_table_fun_future = com_table_fun_all %>% 
  filter(rownames(.) %in% rownames(com_table_future))
  
```


## Veg all
```{r}
# vegNMDS10 = metaMDS(veg_data_match10, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE, autotransform = FALSE)


# dimcheckMDS(veg_data_match10, distance = "bray", k = 10, trymax = 50, autotransform = FALSE)

vegNMDS = metaMDS(com_table, distance = "bray", k = 3, maxit = 999, trymax = 500, wascores = TRUE, autotransform = FALSE)

veg_scores = as.data.frame(scores(vegNMDS))

```

```{r}
soil_veg_scores = envfit(veg_scores, soil_veg, nperm = 999, na.rm = T, choices = c(1:3))
soil_veg_scores_plot = as.data.frame(scores(soil_veg_scores, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(variable = case_when(rowname == "change18_10" ~ "0-10cm % Carbon Change",
                              rowname == "2018_10" ~ "0-10cm % Organic C",
                              rowname == "change18_40" ~ "10-40cm % Carbon Change",
                              rowname == "2018_40" ~ "10-40cm % Organic C"
                              ),
         label = case_when(rowname == "change18_10" ~ "A",
                              rowname == "2018_10" ~ "B",
                              rowname == "change18_40" ~ "C",
                              rowname == "2018_40" ~ "D"))


fun_grp_scores = envfit(veg_scores, com_table_fun_all, nperm = 999, choices = c(1:3))
fun_grp_scores_plot = as.data.frame(scores(fun_grp_scores, display = "vectors")) %>% 
  rownames_to_column() %>% 
  left_join(as.data.frame(fun_grp_scores$vectors$pvals) %>% rownames_to_column(), by = "rowname") %>% 
  filter(`fun_grp_scores$vectors$pvals` < 0.05) %>% 
  mutate(functional_group = case_when(rowname == "SedgesRushes" ~ "Sedges/Rushes",
                                      rowname != "SedgesRushes" ~ rowname))

  # left_join(as.data.frame(fam_16_env10$vectors$pvals) %>% rownames_to_column(), by = "rowname")

veg_scores = as.data.frame(scores(vegNMDS)) %>% 
  rownames_to_column() %>% 
  left_join(soil_veg %>% rownames_to_column(), by = "rowname") 

```


```{r}
soil_veg_scores
fun_grp_scores
```



```{r}
(veg_1_2 = ggplot(veg_scores, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = `2018_10`), size = 3) +
  geom_segment(data = soil_veg_scores_plot, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "black", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text_repel(data = soil_veg_scores_plot, aes(x = NMDS1, y = NMDS2, label = label),
            size = 6, box.padding = 0.25) +
  geom_text_repel(data = fun_grp_scores_plot, mapping = aes(x = NMDS1, y = NMDS2, label = functional_group), color = "blue", size = 6) +
  scale_color_viridis_c(name = "% Organic C 0-10 cm") + 
  theme_bw()) +
  theme(axis.title = element_text(size = 14))

ggsave(here::here("documents/figures/0_10_veg_NMDS_12.png"), dpi = 300, plot = veg_1_2, device = "png", width = 8, height = 6)


```

```{r}
(veg_1_3 = ggplot(veg_scores, aes(NMDS1, NMDS3)) +
  geom_point(aes(color = `2018_10`), size = 3)+
  geom_segment(data = soil_veg_scores_plot, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS3), 
               color = "black", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text_repel(data = soil_veg_scores_plot, aes(x = NMDS1, y = NMDS3, label = label),
            size = 6, box.padding = 0.25) +
  geom_text_repel(data = fun_grp_scores_plot, mapping = aes(x = NMDS1, y = NMDS3, label = functional_group), color = "blue", size = 6) +
  scale_color_viridis_c(name = "% Organic C") +
  theme_bw())  +
  theme(axis.title = element_text(size = 14))

ggsave(here::here("documents/figures/0_10_veg_NMDS_13.png"), dpi = 300, plot = veg_1_3, device = "png", width = 8, height = 6)
```
```{r}
library(ggpubr)

(veg_NMDS_all = ggarrange(veg_1_2, veg_1_3, common.legend = TRUE, legend = "right", ncol = 1)  +
  theme(axis.title = element_text(size = 12)))

ggsave(here::here("documents/figures/veg_NMDS_all.png"), dpi = 300, plot = veg_NMDS_all, device = "png", width = 8, height = 12)
```


```{r}
veg_table_soil = com_table %>% 
  rownames_to_column() %>% 
  left_join(soil_veg %>% rownames_to_column(), by = "rowname") %>% 
  left_join(soil_veg_future, by = c("rowname" = "point_id"))


napu_mod = lm(`2018_10` ~ DACA3, data = veg_table_soil)
summary(napu_mod)

```


## diversity

```{r}
divs = diversity(com_table, index = "shannon") %>% 
  as.data.frame()

colnames(divs) = "diversity18"

divs_21 = diversity(com_table21_change) %>% 
  as.data.frame() 

colnames(divs_21) = "diversity21"

diversity21 = divs_21 %>% 
  rownames_to_column()

diversity_soil = divs %>% 
  select(diversity18) %>% 
  rownames_to_column() %>% 
  left_join(soil_veg %>% rownames_to_column(), by = "rowname") %>% 
  left_join(soil_veg_future, by = c("rowname" = "point_id")) %>% 
  left_join(diversity21, by = "rowname") %>% 
  mutate(diversity_change = diversity21 - diversity18)


div_surf_mod = lm(`2018_10` ~ diversity18, data = diversity_soil)
summary(div_surf_mod)

div_depth_mod = lm(`2018_40` ~ diversity18, data = diversity_soil)
summary(div_depth_mod)

div_surf_change_mod = lm(change18_10 ~ diversity18, data = diversity_soil) #significant
summary(div_surf_change_mod)

div_depth_change_mod = lm(change18_40 ~ diversity18, data = diversity_soil)
summary(div_depth_change_mod)

div_surf_future_mod = lm(surface ~ diversity18, data = diversity_soil)
summary(div_surf_future_mod)

div_depth_future_mod = lm(depth ~ diversity18, data = diversity_soil)
summary(div_depth_future_mod)

div_change_surf = lm(surface ~ diversity_change, data = diversity_soil)
summary(div_change_surf)

div_change_surf = lm(depth ~ diversity_change, data = diversity_soil)
summary(div_change_surf)


```

```{r}
ggplot(diversity_soil, aes(x = diversity18, y = change18_10)) +
  geom_point() +
  geom_smooth(method = "lm")

summary(div_surf_change_mod)
plot(div_surf_change_mod)

```


## functional groups

```{r}
fun_grp_plot = com_table_fun_standard10 %>% 
  rownames_to_column() %>% 
  pivot_longer(-rowname) %>% 
  mutate(fxn = factor(name, levels = rev(c("Annual Grass", "Perennial Grass", "Annual Forb", "Perennial Forb", "Legumes", "SedgesRushes", "ShrubsTrees"))))
  


ggplot(fun_grp_plot, aes(x = rowname, y = value, fill = fxn)) +
  geom_col() +
  scale_fill_manual(name = "Growth Form", values = rev(cal_palette("superbloom3", n = 7, type = "continuous")))

top_species = new_com_table %>% 
  rownames_to_column() %>% 
  pivot_longer(-rowname) %>% 
  group_by(name) %>% 
  summarise(total = sum(value))
```


```{r}
soil_veg_scores

# veg_mod = lm(`2018_10` ~ NMDS1 + NMDS2 + NMDS3, data = veg_scores)
# summary(veg_mod)
# 
```

```{r}

fun_table_soil = com_table_fun_all %>% 
  rownames_to_column() %>% 
  left_join(soil_veg %>% rownames_to_column(), by = "rowname") %>% 
  clean_names() %>% 
  mutate(p_a_grass = perennial_grass/annual_grass)

pg_table = com_table %>% 
  select(NAPU4, PHAQ, DACA3) %>% 
  rownames_to_column() %>% 
  left_join(soil_veg %>% rownames_to_column(), by = "rowname")

phaqmod = lm(`2018_10` ~ PHAQ, data = pg_table)
summary(phaqmod)

dacamod = lm(`2018_10` ~ DACA3, data = pg_table)
summary(dacamod)



pg_mod = lm(x2018_10 ~ perennial_grass, data = fun_table_soil)
summary(pg_mod)

pg_mod40 = lm(x2018_40 ~ perennial_grass, data = fun_table_soil)
summary(pg_mod40)

``` 


## Future veg

```{r}
veg_future_NMDS = metaMDS(com_table_future, distance = "bray", k = 2, maxit = 999, trymax = 500, wascores = TRUE, autotransform = FALSE)

veg_scores_future = as.data.frame(scores(veg_future_NMDS))

```


```{r}
soil_veg_scores_future = envfit(veg_scores_future, soil_veg_future, nperm = 999, na.rm = T, choices = c(1:3))
soil_veg_scores_future_plot = as.data.frame(scores(soil_veg_scores_future, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(variable = case_when(rowname == "change18_10" ~ "0-10cm % Carbon Change",
                              rowname == "2018_10" ~ "0-10cm % Organic C",
                              rowname == "change18_40" ~ "10-40cm % Carbon Change",
                              rowname == "2018_40" ~ "10-40cm % Organic C"
                              ),
         label = case_when(rowname == "surface" ~ "0-10cm",
                              rowname == "depth" ~ "10-40cm"))


fun_grp_scores_future = envfit(veg_scores_future, com_table_fun_future, nperm = 999, choices = c(1:3))
fun_grp_scores_future_plot = as.data.frame(scores(fun_grp_scores_future, display = "vectors")) %>% 
  rownames_to_column() %>% 
  left_join(as.data.frame(fun_grp_scores$vectors$pvals) %>% rownames_to_column(), by = "rowname") %>% 
  filter(`fun_grp_scores$vectors$pvals` < 0.05) %>% 
  mutate(functional_group = case_when(rowname == "SedgesRushes" ~ "Sedges/Rushes",
                                      rowname != "SedgesRushes" ~ rowname))

  # left_join(as.data.frame(fam_16_env10$vectors$pvals) %>% rownames_to_column(), by = "rowname")

veg_scores_future = as.data.frame(scores(veg_future_NMDS)) %>% 
  rownames_to_column() %>% 
  left_join(soil_veg_future %>% rownames_to_column(), by = c("rowname" = "point_id")) 

```

```{r}
(veg_future = ggplot(veg_scores_future, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = surface), size = 3) +
  geom_segment(data = soil_veg_scores_future_plot, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "black", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text_repel(data = soil_veg_scores_future_plot, aes(x = NMDS1, y = NMDS2, label = label),
            size = 6, box.padding = 0.1) +
  geom_text_repel(data = fun_grp_scores_future_plot, mapping = aes(x = NMDS1, y = NMDS2, label = functional_group), color = "blue", size = 6) +
  scale_color_viridis_c(name = "Change 2018-21 \n%SOC 0-10 cm ") + 
  theme_bw()) +
  theme(axis.title = element_text(size = 14))

ggsave(here::here("documents/figures/veg_future_NMDS.png"), dpi = 300, plot = veg_future, device = "png", width = 8, height = 6)


soil_veg_scores_future
```


```{r}
fun_table_soil_future = com_table_fun_future %>% 
  rownames_to_column() %>% 
  left_join(soil_veg_future %>% rownames_to_column(), by = c("rowname" = "point_id")) %>% 
  clean_names() %>% 
  mutate(p_a_grass = perennial_grass/annual_grass)
```

```{r}
mod_future = lm(surface ~ log(p_a_grass), data = fun_table_soil_future)
summary(mod_future)

mod_futured = lm(depth ~ log(p_a_grass), data = fun_table_soil_future)
summary(mod_futured)

plot(mod_futured)

```

```{r}
(p_a_change = ggplot(fun_table_soil_future, aes(x = log(p_a_grass), y = surface)) +
  geom_point() +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() +
  theme(axis.title = element_text(size = 14)) +
  xlab("Perennial Grass:Annual Grass") +
  ylab("% Change SOC 0-10cm 2018-21"))
  
(p_a_change = ggplot(fun_table_soil_future, aes(x = log(p_a_grass), y = depth)) +
  geom_point() +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() +
  theme(axis.title = element_text(size = 14)) +
  xlab("Perennial Grass:Annual Grass") +
  ylab("% Change SOC 0-10cm 2018-21"))

ggsave(here::here("documents/figures/p_a_change.png"), dpi = 300, plot = p_a_change, device = "png", width = 8, height = 6)


```

```{r}
(perennial_change = ggplot(fun_table_soil_future, aes(x = perennial_grass, y = surface)) +
  geom_point() +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() +
  theme(axis.title = element_text(size = 14)) +
  xlab("Perennial Grass Cover") +
  ylab("% Change SOC 0-10cm 2018-21"))

(perennial_change = ggplot(fun_table_soil_future, aes(x = perennial_grass, y = depth)) +
  geom_point() +
  geom_smooth(method = "lm", color = "black") +
  theme_bw() +
  theme(axis.title = element_text(size = 14)) +
  xlab("Perennial Grass Cover") +
  ylab("% Change SOC 0-10cm 2018-21"))

ggsave(here::here("documents/figures/perennial_change.png"), dpi = 300, plot = perennial_change, device = "png", width = 8, height = 6)

```




```{r}
veg_soil_change = veg_point_distance %>% 
  mutate(point_id = str_sub(points, 1, 8)) %>% 
  left_join(change, by = "point_id") %>% 
  select(point_id, distance, depth_cm, change18) %>% 
  pivot_wider(values_from = change18, names_from = depth_cm) %>% 
  rename(surface = `0 to 10`,
         depth = `10 to 40`)

m = lm(surface ~ distance, data = veg_soil_change)
summary(m)



```

```{r}

veg_soil_change18 = veg_point_distance18 %>% 
  mutate(point_id = str_sub(points, 1, 8)) %>% 
  left_join(change, by = "point_id") %>% 
  select(point_id, distance, depth_cm, change18) %>% 
  pivot_wider(values_from = change18, names_from = depth_cm) %>% 
  rename(surface = `0 to 10`,
         depth = `10 to 40`)

m18 = lm(depth ~ distance, data = veg_soil_change18)
summary(m18)

```

```{r}
ggplot(veg_soil_change, aes(x = distance, y = surface)) +
  geom_point() +
  theme_bw() +
  ylab("Change 2018 - 2021")

ggplot(veg_soil_change, aes(x = distance, y = depth)) +
  geom_point() +
  theme_bw() +
  ylab("Change 2018 - 2021")
```


## Veg change matrix


```{r}

```


```{r}



veg_change_com_table = veg_change_com

veg_change_fun = veg_change_com_table %>% 
  rownames_to_column("point_id") %>% 
  pivot_longer(-point_id) %>% 
  left_join(fungrps, by = c("name" = "symbol")) %>% 
  mutate(fun_grp = replace(fun_grp, name == "2GA", "Annual Grass"),
         fun_grp = replace(fun_grp, name == "2FA", "Annual Forb"),
         fun_grp = replace(fun_grp, name == "2GP", "Perennial Grass")) %>% 
  select(-name) %>% 
  group_by(point_id, fun_grp) %>% 
  summarise(fun_change = sum(value)) %>% 
  filter(!is.na(fun_grp)) %>% 
  pivot_wider(point_id, names_from = fun_grp, values_from = fun_change)


fun_table21 = data_21 %>% 
  select(-year) %>% 
  pivot_longer(-point_id) %>% 
  left_join(fungrps, by = c("name" = "symbol")) %>% 
  mutate(fun_grp = replace(fun_grp, name == "2GA", "Annual Grass"),
         fun_grp = replace(fun_grp, name == "2FA", "Annual Forb"),
         fun_grp = replace(fun_grp, name == "2GP", "Perennial Grass")) %>% 
  select(-name) %>% 
  group_by(point_id, fun_grp) %>% 
  summarise(fun_change = sum(value)) %>% 
  filter(!is.na(fun_grp)) %>% 
  pivot_wider(point_id, names_from = fun_grp, values_from = fun_change) %>% 
  column_to_rownames("point_id") %>% 
  decostand(method = "total") %>% 
  clean_names() %>% 
  mutate(p_a_grass = perennial_grass/annual_grass) %>% 
  rownames_to_column("point_id")



           
           

```

```{r}
veg_change_NMDS = metaMDS(veg_change_com_table, distance = "euclidian", k = 2, maxit = 999, trymax = 500, wascores = TRUE, autotransform = T)

veg_scores_change = as.data.frame(scores(veg_change_NMDS))

```

```{r}

soil_veg_scores_change = envfit(veg_scores_change, soil_veg_future, nperm = 999, na.rm = T, choices = c(1:2))
soil_veg_scores_change_plot = as.data.frame(scores(soil_veg_scores_change, display = "vectors")) %>% 
  rownames_to_column() %>% 
  mutate(variable = case_when(rowname == "change18_10" ~ "0-10cm % Carbon Change",
                              rowname == "2018_10" ~ "0-10cm % Organic C",
                              rowname == "change18_40" ~ "10-40cm % Carbon Change",
                              rowname == "2018_40" ~ "10-40cm % Organic C"
                              ),
         label = case_when(rowname == "surface" ~ "0-10cm",
                              rowname == "depth" ~ "10-40cm"))


fun_grp_scores_change = envfit(veg_scores_change, veg_change_fun, nperm = 999, choices = c(1:3))
fun_grp_scores_change_plot = as.data.frame(scores(fun_grp_scores_change, display = "vectors")) %>% 
  rownames_to_column() %>% 
  left_join(as.data.frame(fun_grp_scores$vectors$pvals) %>% rownames_to_column(), by = "rowname") %>% 
  filter(`fun_grp_scores$vectors$pvals` < 0.05) %>%
  mutate(functional_group = case_when(rowname == "SedgesRushes" ~ "Sedges/Rushes",
                                      rowname != "SedgesRushes" ~ rowname))

  # left_join(as.data.frame(fam_16_env10$vectors$pvals) %>% rownames_to_column(), by = "rowname")

veg_scores_change = as.data.frame(scores(veg_change_NMDS)) %>% 
  rownames_to_column() %>% 
  left_join(soil_veg_future %>% rownames_to_column(), by = c("rowname" = "point_id")) 

```

```{r}
(veg_change = ggplot(veg_scores_change, aes(NMDS1, NMDS2)) +
  geom_point(aes(color = surface), size = 3) +
  geom_segment(data = soil_veg_scores_change_plot, aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2), 
               color = "black", 
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text_repel(data = soil_veg_scores_change_plot, aes(x = NMDS1, y = NMDS2, label = label),
            size = 6, box.padding = 0.1) +
  geom_text_repel(data = fun_grp_scores_change_plot, mapping = aes(x = NMDS1, y = NMDS2, label = functional_group), color = "blue", size = 6) +
  scale_color_viridis_c(name = "Change 2018-21 \n%SOC 0-10 cm ") + 
  theme_bw()) +
  theme(axis.title = element_text(size = 14))

soil_veg_scores_change

```

```{r}
p_a_change = tibble(p_a_change = fun_table21$p_a_grass - fun_table_soil_future$p_a_grass)



fun_table_soil_change = veg_change_fun %>% 
  rownames_to_column() %>% 
  left_join(soil_veg_future %>% rownames_to_column(), by = "point_id") %>% 
  clean_names() %>% 
  bind_cols(p_a_change) %>% 
  mutate(p_a_change_transform = log(p_a_change - min(fun_table_soil_change$p_a_change) + 0.01))


pg_c = lm(surface ~ log(perennial_grass + 0.4), data = fun_table_soil_change)
summary(pg_c)

pg_cd = lm(depth ~ (perennial_grass), data = fun_table_soil_change)
summary(pg_cd)

plot(pg_c)
```
```{r}
ggplot(fun_table_soil_change, aes(x = log(perennial_grass + 0.4), y = surface)) + 
  geom_point()
```










# canopy

```{r}
canopy_resample = canopy_sum %>% 
  group_by(point_id) %>% 
  filter(n() > 1)

canopy_total = canopy_sum %>% 
  group_by(point_id, year) %>% 
  rowwise() %>% 
  summarise(total = sum(c_across("BAPI":"RUAR9")))

canopy_total_18 = canopy_total %>% 
  filter(year == "2018")

canopy18 = canopy_sum %>% 
  filter(year == "2018")
```

```{r}
canopy_total_18_soil = canopy_total_18 %>% 
  left_join(change, by = "point_id") %>% 
  filter(point_id != "TOKA-013") %>% 
  select(point_id, total, depth_cm, change18, `2018`, change21) %>% 
  pivot_wider(values_from =change21 , names_from = depth_cm) %>% 
  rename(surface = `0 to 10`,
         depth = `10 to 40`)
```


```{r}
cm = lm(`surface` ~ total, data = canopy_total_18_soil)
summary(cm)


cm = lm(`depth` ~ total, data = canopy_total_18_soil)
summary(cm)
```

```{r}
ggplot(canopy_total_18_soil, aes(x = total, y = depth)) +
  geom_point() +
  scale_color_manual(name = "Depth", values = c("green", "purple")) +
  xlab("Canopy Cover 2018") +
  ylab("% SOC 2018") +
  theme_bw()

ggplot(canopy_total_18_soil, aes(x = total, y = surface)) +
  geom_point() +
  scale_color_manual(name = "Depth", values = c("green", "purple")) +
  xlab("Canopy Cover 2018") +
  ylab("% SOC 2018") +
  theme_bw()
```



```{r}
canopy_change = canopy_cover %>% 
  ungroup() %>% 
  group_by(point_id) %>% 
  filter(n() > 1) %>% 
  summarise(canopy_change = diff(percent_cover))

canopy_change_soil = canopy_change %>% 
  left_join(change, by = "point_id") %>% 
  select(point_id, canopy_change, depth_cm, change21) %>% 
  pivot_wider(values_from = change21, names_from = depth_cm) %>% 
  rename(surface = `0 to 10`,
         depth = `10 to 40`)


ccm = lm(surface ~ canopy_change, data = canopy_change_soil)
summary(ccm)

ccm1 = lm(depth ~ canopy_change, data = canopy_change_soil)
summary(ccm1)

plot(ccm)

```

```{r}
ggplot(canopy_change_soil, aes(x = canopy_change, y = surface)) +
  geom_point() +
  scale_color_manual(name = "Depth", values = c("green", "purple")) +
  xlab("Canopy Cover Change 2018 - 2021") +
  ylab("SOC Change 2021") +
  geom_smooth(method = "lm") +
  theme_bw()


ggplot(canopy_change_soil, aes(x = canopy_change, y = depth)) +
  geom_point() +
  scale_color_manual(name = "Depth", values = c("green", "purple")) +
  xlab("Canopy Cover Change 2018 - 2021") +
  ylab("SOC Change 2021") +
  geom_smooth(method = "lm") +
  theme_bw()
```

```{r}
plot(ccm)
```



```{r}
vrda = rda(com_table_fun_standard10)
summary(vrda)

vrda_scores = as.data.frame(scores(vrda, display = "sites")) %>% 
  rownames_to_column() %>% 
  left_join(soil_veg_match10 %>% rownames_to_column(), by = "rowname") 
  


vrda_mod = lm(change18 ~ PC1 + PC2, data = vrda_scores)
summary(vrda_mod)
```

# Fungal analyses



## full community
```{r}

ASV_tab = asv_table_standard %>% 
  filter(rowSums(.) > 0)

ASV_ord_all = metaMDS(com_table, distance = "bray", k = 3, maxit = 999, trymax = 500, wascores = TRUE, autotransform = FALSE)

```





























