---
title: "Variance Partitioning"
author: "Jacob Weverka"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
```

Read in cleanup from other files
```{r results="hide"}
knitr::knit(here("code/soil_data_cleanup.Rmd"))
knitr::knit(here("code/veg_data_cleanup.Rmd"))
knitr::knit(here("code/16S_cleanup.Rmd"))
knitr::knit(here("code/fungal_cleanup.Rmd"))
```


## Data cleanup

Trim data to intersection of microbial data and change data
```{r}

change[41:80,5:7] = change[1:40,5:7]

change_data = change %>% 
  mutate(id = case_when(depth_cm == "10 to 40" ~ paste(point_id, "-40", sep = ""),
                        depth_cm == "0 to 10" ~ paste(point_id, "-10", sep = ""))
         ) %>% 
  # filter(id %in% rownames(otu_standard),
  #        id != "TOKA-099-10") %>% 
  column_to_rownames("id")



change_data10 = change_data %>% 
  filter(depth_cm == "0 to 10") %>% 
  as.data.frame() %>% 
  filter(rownames(.) %in% rownames(new_com_table10)) %>% 
  filter(rownames(.) %in% rownames(otu_standard))

change_data40 = change_data %>% 
  filter(depth_cm == "10 to 40") %>% 
  as.data.frame() %>% 
  filter(rownames(.) %in% rownames(new_com_table40)) %>% 
  filter(rownames(.) %in% rownames(otu_standard))

asv_10_match = asv_10 %>% 
  filter(rownames(.) %in% rownames(change_data10))

asv_40_match = asv_40 %>% 
  filter(rownames(.) %in% rownames(change_data40))

all_change = bind_rows(change_data10, change_data40)



c_change = toka_c_points %>% 
  filter(rownames(.) %in% rownames(otu_standard)) %>% 
  filter(`2021` != 0)


otu_standard_match10 = otu_standard %>% 
  filter(rownames(.) %in% rownames(change_data10)) %>% 
  as.data.frame()

otu_standard_match40 = otu_standard %>% 
  filter(rownames(.) %in% rownames(change_data40)) %>% 
  as.data.frame()

otu_standard_match = bind_rows(otu_standard_match10, otu_standard_match40)

family_standard_match10 = family_standard %>%
  filter(rownames(.) %in% rownames(change_data10))

family_standard_match40 = family_standard %>%
  filter(rownames(.) %in% rownames(change_data40))

family_standard_match = bind_rows(family_standard_match10, family_standard_match40)

veg_data_match10 = new_com_table10 %>%
  filter(rownames(.) %in% rownames(change_data10))

veg_data_match40 = new_com_table40 %>%
  filter(rownames(.) %in% rownames(change_data40))
# 
# class_standard_match = class_standard %>% 
#   filter(rownames(.) %in% rownames(change_data))
# 
# otu_standard3 = otu_standard %>% 
#   filter(rownames(.) %in% rownames(c_change))

# family_standard3 = family_standard %>% 
#   filter(rownames(.) %in% rownames(c_change))
# 
# class_standard3 = class_standard %>% 
#   filter(rownames(.) %in% rownames(change_data))

#separate into surface and depth
change_mat10 = change %>% 
  filter(depth_cm == "0 to 10") %>% 
  column_to_rownames("point_id") %>% 
  filter(rownames(.) %in% canopy_sum$point_id) %>% 
  select(n_change, `2018`, n_2018, cn_change, change18, c_n_ratio, sand, silt, clay)

change_mat40 = change %>% 
  filter(depth_cm == "10 to 40") %>% 
  column_to_rownames("point_id") %>% 
  filter(rownames(.) %in% canopy_sum$point_id) %>% 
  select(n_change, `2018`, n_2018, cn_change, change18, c_n_ratio, sand, silt, clay)

#clean canopy data and select
canopy_mat10 = canopy_sum %>% 
  filter(year == 2018) %>% 
  mutate(point_id = paste(point_id, "-10", sep = "")) %>% 
  column_to_rownames("point_id") %>% 
  select(-year) %>% 
  mutate(DIAU = DIAU + DIAUA) %>% 
  select(-DIAUA)%>% 
  filter(rownames(.) %in% rownames(change_data10)) %>% 
  as.data.frame()

canopy_mat40 = canopy_sum %>% 
  filter(year == 2018) %>% 
  mutate(point_id = paste(point_id, "-40", sep = "")) %>% 
  column_to_rownames("point_id") %>% 
  select(-year) %>% 
  mutate(DIAU = DIAU + DIAUA) %>% 
  select(-DIAUA) %>% 
  filter(rownames(.) %in% rownames(change_data40)) %>% 
  as.data.frame()

z = bind_cols(change_data10, canopy_mat10, otu_standard_match10)

# c_10 = z[,c(6,10)]
c_10 = z %>% 
  select(`2018`, change18)
# rownames(c_10) = rownames(z)
# colnames(c_10) = "c_2018"
soil10 = z %>% 
  select(sand, silt, clay)

c_40 = change_data40 %>% 
  select(`2018`, change18)

soil40 = change_data40 %>% 
  select(sand, silt, clay)

# change_data_vp10 = change_data10 %>% 

```



## Analysis
2 Strategies: First, try to do unconstrained PCA on each data set, keep enough axes such that >= 80% of variation is explained in c stock or c_change. Then use axes in variance partitioning

Second strategy: try constrained pca on each data set, constrain with c_change or c stock

### 16S NMDS

NMDS

```{r}

set.seed(44)

otuNMDS10 = metaMDS(otu_standard_match10, distance = "bray", k = 2, try = 50, trymax = 100)
otuNMDS10_scores = scores(otuNMDS10) %>% 
  as.data.frame()

otuNMDS40 = metaMDS(otu_standard_match40, distance = "bray", k = 2, try = 50, trymax = 100)
otuNMDS40_scores = scores(otuNMDS40) %>% 
  as.data.frame()


# 
# 
# plot(otuNMDS_scores10$NMDS1, otuNMDS_scores10$NMDS2, xlim = c(-1, 1), ylim = c(-1, 1))
# points(otuNMDS_scores40$NMDS1, otuNMDS_scores40$NMDS2, col = "red")

```


```{r}

# phylumNMDS = metaMDS(family_standard, distance = "bray", k = 2, try = 50, trymax = 100)
# phylumNMDS_scores = scores(phylumNMDS)
# 
# phylumNMDS_scores10 = phylumNMDS_scores %>%
#   as.data.frame() %>%
#   filter(rownames(.) %in% rownames(change_data10))
# 
# phylumNMDS_scores40 = phylumNMDS_scores %>%
#   as.data.frame() %>%
#   filter(rownames(.) %in% rownames(change_data40))
# 
# 
# plot(phylumNMDS_scores10$NMDS1, phylumNMDS_scores10$NMDS2, xlim = c(-1, 1), ylim = c(-1, 1))
# points(phylumNMDS_scores40$NMDS1, phylumNMDS_scores40$NMDS2, col = "red")
# 

```

```{r}
# canopydist = vegdist(canopy_mat40, method = "bray")
# canopypcoa = cmdscale(canopydist, eig = TRUE, k = 6)
# canopy.var.per = (canopypcoa$eig/sum(canopypcoa$eig))*100
# canopypcoa
# 
# canopy40PCOA_scores = scores(canopypcoa, choices = c(1:6)) %>% 
#   as.data.frame()
# 
# canopy10PCOA_scores = canopy40PCOA_scores %>% 
#   rownames_to_column() %>% 
#   mutate(rowname = str_replace(rowname, "-40", "-10")) %>% 
#   column_to_rownames("rowname") %>% 
#   filter(rownames(.) %in% rownames(canopy_mat10))
# 


```

### Canopy NMDS

```{r}

canopyNMDS40 = metaMDS(canopy_mat40, distance = "bray", k = 2, try = 50, trymax = 100)

canopyNMDS_scores40 = scores(canopyNMDS40) %>% as.data.frame()

canopyNMDS_scores10 = canopyNMDS_scores40  %>%
  filter(rownames(.) != "TOKA-099-40") %>%
  rownames_to_column() %>% 
  mutate(rowname = str_replace(rowname, "-40", "-10")) %>%
  column_to_rownames("rowname")



```

```{r}
# veg_dist = vegdist(veg_data_match40, method = "bray")
# veg40pcoa = cmdscale(veg_dist, eig = TRUE, k = 6)
# veg.var.per = (veg40pcoa$eig/sum(veg40pcoa$eig))*100
# veg40pcoa
# 
# veg40PCOA_scores = scores(veg40pcoa, choices = c(1:6)) %>% 
#   as.data.frame()
# 
# veg10PCOA_scores = veg40PCOA_scores %>% 
#   rownames_to_column() %>% 
#   mutate(rowname = str_replace(rowname, "-40", "-10")) %>% 
#   column_to_rownames("rowname") %>% 
#   filter(rownames(.) %in% rownames(canopy_mat10))

```


### Herbaceous NMDS

```{r}

vegNMDS40 = metaMDS(veg_data_match40, distance = "bray", k = 2, try = 50, trymax = 100)

vegNMDS_scores40 = scores(vegNMDS40) %>% as.data.frame()

vegNMDS_scores10 = vegNMDS_scores40 %>%
  filter(rownames(.) != "TOKA-099-40") %>%
  rownames_to_column() %>% 
  mutate(rowname = str_replace(rowname, "-40", "-10")) %>%
  column_to_rownames("rowname")


```
```{r}

```



```{r}
all_veg10 = bind_cols(canopyNMDS_scores10, vegNMDS_scores10)
all_veg40 = bind_cols(canopyNMDS_scores40, vegNMDS_scores40)

# all_veg10pcoa = bind_cols(canopy10PCOA_scores, veg10PCOA_scores)
# all_veg40pcoa = bind_cols(canopy40PCOA_scores, veg40PCOA_scores)

```

### Fungal NMDS
```{r}

# fungal_NMDS = metaMDS(asv_all, distance = "bray", k = 2)
# 
# fungal_scores10 = scores(fungal_NMDS) %>% 
#   as.data.frame() %>% 
#   filter(rownames(.) %in% rownames(asv_10))
# 
# fungal_scores40 = scores(fungal_NMDS) %>% 
#   as.data.frame() %>% 
#   filter(rownames(.) %in% rownames(asv_40))



fungal_NMDS10 = metaMDS(asv_10_match, distance = "bray", k = 2, try = 50, trymax = 100)

fungal_scores10 = scores(fungal_NMDS10) %>%
  as.data.frame()

fungal_NMDS40 = metaMDS(asv_40_match, distance = "bray", k = 2, try = 50, trymax = 100)

fungal_scores40 = scores(fungal_NMDS40) %>%
  as.data.frame()
# 
# plot(fungal_scores10$NMDS1, fungal_scores10$NMDS2, xlim = c(-1, 1), ylim = c(-1, 1))
# points(fungal_scores40$NMDS1, fungal_scores40$NMDS2, col = "red")

```


```{r}
c_stock_10 = varpart(z$`2018`, all_veg10, otuNMDS10_scores, fungal_scores10)
# c_stock_10

plot(c_stock_10, bg = c("green", "brown", "purple"), Xnames = c("Plants", "Microbes", "Fungus"))




```

```{r}
c_stock_40 = varpart(c_40$`2018`, all_veg40, otuNMDS40_scores, fungal_scores40)

plot(c_stock_40,  bg = c("green", "brown", "purple"),  Xnames = c("Plants", "Microbes", "Fungus" ))


```

### C stocks figure

```{r}


png(here("documents/figures/varpar_c_stock.png"), width = 1000, height = 500)

par(mfrow = c(1,2))

plot(c_stock_10, bg = c("green", "brown", "purple"), Xnames = c("Plants", "Bacteria", "Fungus"), main = "0-10 cm")
title(main = "0-10 cm")
plot(c_stock_40,  bg = c("green", "brown", "purple"),  Xnames = c("Plants", "Bacteria", "Fungus"), main = "10-40 cm")
title(main = "10-40 cm")

dev.off()



```






```{r}
c_change_10 = varpart(c_10$change18, all_veg10, otuNMDS10_scores, fungal_scores10)
# c_stock_10

plot(c_change_10, bg = c("green", "brown", "purple"), Xnames = c("Plants", "16S", "Fungus" ))

# c_change_10family = varpart(c_10$change18, all_veg10, phylumNMDS_scores10, soil10)
# # c_stock_10
# 
# plot(c_change_10family, digits= 2,  bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ))
# 

# showvarparts(3)


```

```{r}
c_change_10_veg = varpart(c_10$change18, canopyNMDS_scores10, vegNMDS_scores10)

plot(c_change_10_veg)
```

```{r}
c_change_40 = varpart(c_40$change18, all_veg40, otuNMDS40_scores, fungal_scores40)
# c_stock_10

plot(c_change_40, bg = c("green",  "purple", "brown"), Xnames = c("Plants", "16S", "Fungus" ))
# 
# c_change_40family = varpart(c_40$change18, all_veg40, phylumNMDS_scores40, soil40)
# # c_stock_10
# 
# plot(c_change_40family, bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ))


```

### change 18 figure

```{r}
png(here("documents/figures/varpart_change18.png"), width = 1000, height = 500)

par(mfrow = c(1,2))

plot(c_change_10, bg = c("green", "brown", "purple"), Xnames = c("Plants", "Bacteria", "Fungus" ))
title(main = "0-10 cm")
plot(c_change_40, bg = c("green",  "purple", "brown"), Xnames = c("Plants", "Bacteria", "Fungus" ))
title(main = "10-40 cm")

dev.off()


```


```{r}
otu_scores_change10 = otuNMDS10_scores %>% 
  filter(rownames(.) %in% rownames(c_change))

otu_scores_change40 = otuNMDS40_scores %>% 
  filter(rownames(.) %in% rownames(c_change))

c_change_surface = c_change %>% 
  filter(rownames(.) %in% rownames(otu_scores_change10))

c_change_deep = c_change %>% 
  filter(rownames(.) %in% rownames(otu_scores_change40))

# phylum_scores_change10 = phylumNMDS_scores10 %>% 
#   filter(rownames(.) %in% rownames(c_change))
# 
# phylum_scores_change40 = phylumNMDS_scores40 %>% 
#   filter(rownames(.) %in% rownames(c_change))

veg_scores_change10 = vegNMDS_scores10 %>% 
  filter(rownames(.) %in% rownames(c_change))

veg_scores_change40 = vegNMDS_scores40 %>% 
  filter(rownames(.) %in% rownames(c_change))

canopy_scores_change10 = canopyNMDS_scores10 %>% 
  filter(rownames(.) %in% rownames(c_change))

canopy_scores_change40 = canopyNMDS_scores40 %>% 
  filter(rownames(.) %in% rownames(c_change))

all_veg_scores_change10 = veg_scores_change10 %>% 
  bind_cols(canopy_scores_change10)

all_veg_scores_change40 = veg_scores_change40 %>% 
  bind_cols(canopy_scores_change40)

soil40_change = soil40 %>% 
  filter(rownames(.) %in% rownames(c_change))

soil10_change = soil10 %>% 
  filter(rownames(.) %in% rownames(c_change))

fungal10_change = fungal_scores10 %>% 
  filter(rownames(.) %in% rownames(c_change))

fungal40_change = fungal_scores40 %>% 
  filter(rownames(.) %in% rownames(c_change))


```

```{r}

c_change_40_21 = varpart(c_change_deep$change21, all_veg_scores_change40, otu_scores_change40, fungal40_change)
# c_stock_10
plot(c_change_40_21, bg = c("green", "brown", "purple"), Xnames = c("Plants", "16S", "Fungus" ))

c_change_40_21_x = varpart(c_change_deep$change21, otu_scores_change40, fungal40_change)
plot(c_change_40_21_x)


# c_change_40family_21 = varpart(c_change_deep$change21, all_veg_scores_change40, phylum_scores_change40, soil40_change)
# # c_stock_10
# plot(c_change_40family_21, bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ))

```




```{r}

c_change_10_21 = varpart(c_change_surface$change21, veg_scores_change10, otu_scores_change10, fungal10_change)
# c_stock_10
plot(c_change_10_21, bg = c("green", "brown", "purple"), Xnames = c("Plants", "16S", "Fungus" ))

# 
# c_change_10family_21 = varpart(c_change_surface$change21, veg_scores_change10, phylum_scores_change10, soil10_change)
# # c_stock_10
# plot(c_change_10family_21, bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ))

```
### future plot

```{r}

png(here("documents/figures/varpart_change21.png"), width = 1000, height = 500)

par(mfrow = c(1,2))

plot(c_change_10_21, bg = c("green", "brown", "purple"), Xnames = c("Plants", "Bacteria", "Fungus" ))
title(main = "0-10 cm")
plot(c_change_40_21, bg = c("green", "brown", "purple"), Xnames = c("Plants", "Bacteria", "Fungus" ))
title(main = "10-40 cm")

dev.off()


```


```{r}
# otu_data10 = bind_cols(c_10, otuNMDS_scores10)
# otu_model_10 = lm(c_2018 ~ NMDS1 + NMDS2, data = otu_data10)
# summary(otu_model_10)
# 
# 
# 
# otu_data40 = bind_cols(c_40, otuNMDS_scores40)
# otu_model_40 = lm(c_2018 ~ NMDS1 + NMDS2, data = otu_data40)
# summary(otu_model_40)

```

### All Plots

```{r}

png(here("documents/figures/varpart_all.png"), width = 1200, height = 1600)


par(mfrow = c(3,2))

#stocks
plot(c_stock_10, bg = c("green", "brown", "purple"), Xnames = c("Plants", "Bacteria", "Fungi"), cex = 2.5, cex.sub = 1.5, cex.lab = 1.5, id.size = 3)
title(main = "A. %Organic C, 0-10 cm", cex.main = 2.5)
plot(c_stock_40,  bg = c("green", "brown", "purple"),  Xnames = c("Plants", "Bacteria", "Fungi"), cex = 2.5, cex.sub = 1.5, cex.lab = 1.5, id.size = 3)
title(main = "B. %Organic C, 10-40 cm", cex.main = 2.5)

#change 18
plot(c_change_10, bg = c("green", "brown", "purple"), Xnames = c("Plants", "Bacteria", "Fungi" ), cex = 2.5, cex.sub = 1.5, cex.lab = 1.5, id.size = 3)
title(main = "C. Change 2015-2018 %Organic C, 0-10 cm", cex.main = 2.5)
plot(c_change_40, bg = c("green",  "purple", "brown"), Xnames = c("Plants", "Bacteria", "Fungi" ), cex = 2.5, cex.sub = 1.5, cex.lab = 1.5, id.size = 3)
title(main = "D. Change 2015-2018 %Organic C, 10-40 cm", cex.main = 2.5)

#change 21
plot(c_change_10_21, bg = c("green", "brown", "purple"), Xnames = c("Plants", "Bacteria", "Fungi" ), cex = 2.5, cex.sub = 1.5, cex.lab = 1.5, id.size = 3)
title(main = "E. Change 2018-2021 %Organic C, 0-10 cm", cex.main = 2.5)
plot(c_change_40_21, bg = c("green", "brown", "purple"), Xnames = c("Plants", "Bacteria", "Fungi" ), cex = 2.5, cex.sub = 1.5, cex.lab = 1.5, id.size = 3)
title(main = "F. Change 2018-2021 %Organic C, 10-40 cm", cex.main = 2.5)


dev.off()
```

