---
title: "Variance Partitioning"
author: "Jacob Weverka"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
```

Read in cleanup from other files
```{r results="hide"}
knitr::knit(here("code/soil_data_cleanup.Rmd"))
knitr::knit(here("code/veg_data_cleanup.Rmd"))
knitr::knit(here("code/16S_cleanup.Rmd"))
```



Trim data to intersection of microbial data and change data
```{r}

change[41:80,5:7] = change[1:40,5:7]

change_data = change %>% 
  mutate(id = case_when(depth_cm == "10 to 40" ~ paste(point_id, "-40", sep = ""),
                        depth_cm == "0 to 10" ~ paste(point_id, "-10", sep = ""))
         ) %>% 
  filter(id %in% rownames(otu_standard),
         id != "TOKA-099-10") %>% 
  column_to_rownames("id")



change_data10 = change_data %>% 
  filter(depth_cm == "0 to 10") %>% 
  as.data.frame() %>% 
  filter(rownames(.) %in% rownames(new_com_table10))

change_data40 = change_data %>% 
  filter(depth_cm == "10 to 40") %>% 
  as.data.frame() %>% 
  filter(rownames(.) %in% rownames(new_com_table40))

all_change = bind_rows(change_data10, change_data40)



c_change = toka_c_points %>% 
  filter(rownames(.) %in% rownames(otu_standard)) %>% 
  filter(`2021` != 0)


otu_standard_match10 = otu_standard %>% 
  filter(rownames(.) %in% rownames(change_data10)) %>% 
  as.data.frame()

otu_standard_match40 = otu_standard %>% 
  filter(rownames(.) %in% rownames(change_data40)) %>% 
  as.data.frame()

otu_standard_match = bind_rows(otu_standard_match10, otu_standard_match40)

family_standard_match10 = family_standard %>%
  filter(rownames(.) %in% rownames(change_data10))

family_standard_match40 = family_standard %>%
  filter(rownames(.) %in% rownames(change_data40))

family_standard_match = bind_rows(family_standard_match10, family_standard_match40)

veg_data_match10 = new_com_table10 %>%
  filter(rownames(.) %in% rownames(change_data10))

veg_data_match40 = new_com_table40 %>%
  filter(rownames(.) %in% rownames(change_data40))
# 
# class_standard_match = class_standard %>% 
#   filter(rownames(.) %in% rownames(change_data))
# 
# otu_standard3 = otu_standard %>% 
#   filter(rownames(.) %in% rownames(c_change))

# family_standard3 = family_standard %>% 
#   filter(rownames(.) %in% rownames(c_change))
# 
# class_standard3 = class_standard %>% 
#   filter(rownames(.) %in% rownames(change_data))

#separate into surface and depth
change_mat10 = change %>% 
  filter(depth_cm == "0 to 10") %>% 
  column_to_rownames("point_id") %>% 
  filter(rownames(.) %in% canopy_sum$point_id) %>% 
  select(n_change, `2018`, n_2018, cn_change, change18, c_n_ratio, sand, silt, clay)

change_mat40 = change %>% 
  filter(depth_cm == "10 to 40") %>% 
  column_to_rownames("point_id") %>% 
  filter(rownames(.) %in% canopy_sum$point_id) %>% 
  select(n_change, `2018`, n_2018, cn_change, change18, c_n_ratio, sand, silt, clay)

#clean canopy data and select
canopy_mat10 = canopy_sum %>% 
  mutate(point_id = paste(point_id, "-10", sep = "")) %>% 
  column_to_rownames("point_id") %>% 
  select(-year) %>% 
  mutate(DIAU = DIAU + DIAUA) %>% 
  select(-DIAUA)%>% 
  filter(rownames(.) %in% rownames(change_data10)) %>% 
  as.data.frame()

canopy_mat40 = canopy_sum %>% 
  mutate(point_id = paste(point_id, "-40", sep = "")) %>% 
  column_to_rownames("point_id") %>% 
  select(-year) %>% 
  mutate(DIAU = DIAU + DIAUA) %>% 
  select(-DIAUA) %>% 
  filter(rownames(.) %in% rownames(change_data40)) %>% 
  as.data.frame()

z = bind_cols(change_data10, canopy_mat10, otu_standard_match10)

# c_10 = z[,c(6,10)]
c_10 = z %>% 
  select(`2018`, change18)
# rownames(c_10) = rownames(z)
# colnames(c_10) = "c_2018"
soil10 = z %>% 
  select(sand, silt, clay)

c_40 = change_data40 %>% 
  select(`2018`, change18)

soil40 = change_data40 %>% 
  select(sand, silt, clay)

# change_data_vp10 = change_data10 %>% 

```


```{r}
# #remove correlated otu data
# 
# explanatory = bind_cols(canopy10, otu10)
# 
# otu_cor = cor(explanatory)
# otu_cor[lower.tri(otu_cor)] <- 0
# diag(otu_cor) = 0
# 
# x = apply(otu_cor, 2, function(x) any(x > 0.7))
# 
# # m = otu_cor %>% as.data.frame()
# 
# data.new <- explanatory[,!x]
# 
# otu10_reduced = data.new[,8:ncol(data.new)]

```

## Analysis
2 Strategies: First, try to do unconstrained PCA on each data set, keep enough axes such that >= 80% of variation is explained in c stock or c_change. Then use axes in variance partitioning

Second strategy: try constrained pca on each data set, constrain with c_change or c stock

#### Strategy 1

PCOA

```{r}

# otudist = vegdist(otu_standard_match, method = "bray")
# otupcoa = cmdscale(otudist, eig = TRUE, k = 3)
# otu.var.per = (otupcoa$eig/sum(otupcoa$eig))*100
# otupcoa
# otuPCOA_scores = scores(otupcoa, choices = c(1:3)) %>% 
#   as.data.frame()
# 
# otuPCOA_scores40 = otuPCOA_scores %>% 
#   filter(rownames(.) %in% rownames(otu_standard_match40))
# 
# otuPCOA_scores10 = otuPCOA_scores %>% 
#   filter(rownames(.) %in% rownames(otu_standard_match10))
# 
# otu_PCOA_w_data = otuPCOA_scores %>% 
#   as.data.frame() %>% 
#   bind_cols(all_change)
# 
# ggplot(otu_PCOA_w_data, aes(x = Dim1, y = Dim3, color = depth_cm)) + 
#   geom_point()

```

```{r}
# 
# otu10dist = vegdist(otu10, method = "bray")
# otu10pcoa = cmdscale(otu10dist, eig = TRUE)
# otu.var.per = (otu10pcoa$eig/sum(otu10pcoa$eig))*100
# otu10pcoa
# otu10_scores = otu10pcoa$vectors[,1:3]
# 
# ordiplot (otu10pcoa, display = 'sites', type = 'text')

```

NMDS

```{r}

set.seed(44)

otuNMDS = metaMDS(otu_standard, distance = "bray", k = 2, try = 50, trymax = 100)
otuNMDS_scores = scores(otuNMDS)

otuNMDS_scores10 = otuNMDS_scores %>%
  as.data.frame() %>%
  filter(rownames(.) %in% rownames(change_data10))

otuNMDS_scores40 = otuNMDS_scores %>%
  as.data.frame() %>%
  filter(rownames(.) %in% rownames(change_data40))



plot(otuNMDS_scores10$NMDS1, otuNMDS_scores10$NMDS2, xlim = c(-1, 1), ylim = c(-1, 1))
points(otuNMDS_scores40$NMDS1, otuNMDS_scores40$NMDS2, col = "red")

```




```{r}

# familydist = vegdist(family_standard_match, method = "bray")
# familypcoa = cmdscale(familydist, eig = TRUE, k = 3)
# family.var.per = (familypcoa$eig/sum(familypcoa$eig))*100
# familypcoa
# familyPCOA_scores = scores(familypcoa, choices = c(1:3)) %>% 
#   as.data.frame()
# 
# familyPCOA_scores40 = familyPCOA_scores %>% 
#   filter(rownames(.) %in% rownames(family_standard_match40))
# 
# familyPCOA_scores10 = familyPCOA_scores %>% 
#   filter(rownames(.) %in% rownames(family_standard_match10))
# 
# 
# family_PCOA_w_data = familyPCOA_scores %>% 
#   as.data.frame() %>% 
#   bind_cols(all_change)
# 
# ggplot(family_PCOA_w_data, aes(x = Dim1, y = Dim2, color = depth_cm)) + 
#   geom_point()

```

```{r}

phylumNMDS = metaMDS(family_standard, distance = "bray", k = 2, try = 50, trymax = 100)
phylumNMDS_scores = scores(phylumNMDS)

phylumNMDS_scores10 = phylumNMDS_scores %>%
  as.data.frame() %>%
  filter(rownames(.) %in% rownames(change_data10))

phylumNMDS_scores40 = phylumNMDS_scores %>%
  as.data.frame() %>%
  filter(rownames(.) %in% rownames(change_data40))


plot(phylumNMDS_scores10$NMDS1, phylumNMDS_scores10$NMDS2, xlim = c(-1, 1), ylim = c(-1, 1))
points(phylumNMDS_scores40$NMDS1, phylumNMDS_scores40$NMDS2, col = "red")


```

```{r}
# canopydist = vegdist(canopy_mat40, method = "bray")
# canopypcoa = cmdscale(canopydist, eig = TRUE, k = 6)
# canopy.var.per = (canopypcoa$eig/sum(canopypcoa$eig))*100
# canopypcoa
# 
# canopy40PCOA_scores = scores(canopypcoa, choices = c(1:6)) %>% 
#   as.data.frame()
# 
# canopy10PCOA_scores = canopy40PCOA_scores %>% 
#   rownames_to_column() %>% 
#   mutate(rowname = str_replace(rowname, "-40", "-10")) %>% 
#   column_to_rownames("rowname") %>% 
#   filter(rownames(.) %in% rownames(canopy_mat10))
# 


```


```{r}

canopyNMDS40 = metaMDS(canopy_mat40, distance = "bray", k = 2, try = 50, trymax = 100)

canopyNMDS_scores40 = scores(canopyNMDS40) %>% as.data.frame()

canopyNMDS_scores10 = canopyNMDS_scores40  %>%
  filter(rownames(.) != "TOKA-099-40") %>%
  rownames_to_column() %>% 
  mutate(rowname = str_replace(rowname, "-40", "-10")) %>%
  column_to_rownames("rowname")



```

```{r}
# veg_dist = vegdist(veg_data_match40, method = "bray")
# veg40pcoa = cmdscale(veg_dist, eig = TRUE, k = 6)
# veg.var.per = (veg40pcoa$eig/sum(veg40pcoa$eig))*100
# veg40pcoa
# 
# veg40PCOA_scores = scores(veg40pcoa, choices = c(1:6)) %>% 
#   as.data.frame()
# 
# veg10PCOA_scores = veg40PCOA_scores %>% 
#   rownames_to_column() %>% 
#   mutate(rowname = str_replace(rowname, "-40", "-10")) %>% 
#   column_to_rownames("rowname") %>% 
#   filter(rownames(.) %in% rownames(canopy_mat10))

```


```{r}

vegNMDS40 = metaMDS(veg_data_match40, distance = "bray", k = 2, try = 50, trymax = 100)

vegNMDS_scores40 = scores(vegNMDS40) %>% as.data.frame()

vegNMDS_scores10 = vegNMDS_scores40 %>%
  filter(rownames(.) != "TOKA-099-40") %>%
  rownames_to_column() %>% 
  mutate(rowname = str_replace(rowname, "-40", "-10")) %>%
  column_to_rownames("rowname")


```

```{r}
all_veg10 = bind_cols(canopyNMDS_scores10, vegNMDS_scores10)
all_veg40 = bind_cols(canopyNMDS_scores40, vegNMDS_scores40)

# all_veg10pcoa = bind_cols(canopy10PCOA_scores, veg10PCOA_scores)
# all_veg40pcoa = bind_cols(canopy40PCOA_scores, veg40PCOA_scores)

```


```{r}
c_stock_10 = varpart(z$`2018`, all_veg10, otuNMDS_scores10, soil10)
# c_stock_10

plot(c_stock_10, bg = c("green", "brown", "purple"), Xnames = c("Plants", "16S", "Soil Texture" ), cutoff = -0.5)

c_stock_10family = varpart(c_10$`2018`, all_veg10, phylumNMDS_scores10, soil10)
# c_stock_10

plot(c_stock_10family, bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ), cutoff = -0.5)



```


```{r}
c_stock_10_veg = varpart(c_10$`2018`, canopyNMDS_scores10, vegNMDS_scores10)

plot(c_stock_10_veg)
```

```{r}
c_stock_40 = varpart(c_40$`2018`, all_veg40, otuNMDS_scores40, soil40)

c_stock_40family = varpart(c_40$`2018`, all_veg40, phylumNMDS_scores40, soil40)
# c_stock_40



plot(c_stock_40,  bg = c("green", "brown", "purple"),  Xnames = c("Plants", "16S", "Soil Texture" ), cutoff = -0.5)
plot(c_stock_40family,  bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ), cutoff = -0.5)



```




```{r}
c_stock_40_veg = varpart(c_40$`2018`, canopyNMDS_scores40, vegNMDS_scores40)

plot(c_stock_40_veg)
```

```{r}
c_change_10 = varpart(c_10$change18, all_veg10, otuNMDS_scores10, soil10)
# c_stock_10

plot(c_change_10, bg = c("green", "brown", "purple"), Xnames = c("Plants", "16S", "Soil Texture" ), cutoff = -0.5)

c_change_10family = varpart(c_10$change18, all_veg10, phylumNMDS_scores10, soil10)
# c_stock_10

plot(c_change_10family, digits= 2,  bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ))


# showvarparts(3)


```

```{r}
c_change_10_veg = varpart(c_10$change18, canopyNMDS_scores10, vegNMDS_scores10)

plot(c_change_10_veg)
```

```{r}
c_change_40 = varpart(c_40$change18, all_veg40, otuNMDS_scores40, soil40)
# c_stock_10

plot(c_change_40, bg = c("green", "brown", "purple"), Xnames = c("Plants", "16S", "Soil Texture" ), cutoff = -0.5)

c_change_40family = varpart(c_40$change18, all_veg40, phylumNMDS_scores40, soil40)
# c_stock_10

plot(c_change_40family, bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ))


```


```{r}
c_change_40_veg = varpart(c_40$change18, canopyNMDS_scores40, vegNMDS_scores40)

plot(c_change_40_veg)
```

```{r}
otu_scores_change10 = otuNMDS_scores10 %>% 
  filter(rownames(.) %in% rownames(c_change))

otu_scores_change40 = otuNMDS_scores40 %>% 
  filter(rownames(.) %in% rownames(c_change))

c_change_surface = c_change %>% 
  filter(rownames(.) %in% rownames(otu_scores_change10))

c_change_deep = c_change %>% 
  filter(rownames(.) %in% rownames(otu_scores_change40))

phylum_scores_change10 = phylumNMDS_scores10 %>% 
  filter(rownames(.) %in% rownames(c_change))

phylum_scores_change40 = phylumNMDS_scores40 %>% 
  filter(rownames(.) %in% rownames(c_change))

veg_scores_change10 = vegNMDS_scores10 %>% 
  filter(rownames(.) %in% rownames(c_change))

veg_scores_change40 = vegNMDS_scores40 %>% 
  filter(rownames(.) %in% rownames(c_change))

canopy_scores_change10 = canopyNMDS_scores10 %>% 
  filter(rownames(.) %in% rownames(c_change))

canopy_scores_change40 = canopyNMDS_scores40 %>% 
  filter(rownames(.) %in% rownames(c_change))

all_veg_scores_change10 = veg_scores_change10 %>% 
  bind_cols(canopy_scores_change10)

all_veg_scores_change40 = veg_scores_change40 %>% 
  bind_cols(canopy_scores_change40)

soil40_change = soil40 %>% 
  filter(rownames(.) %in% rownames(c_change))

soil10_change = soil10 %>% 
  filter(rownames(.) %in% rownames(c_change))

```

```{r}

c_change_40_21 = varpart(c_change_deep$change21, all_veg_scores_change40, otu_scores_change40, soil40_change)
# c_stock_10
plot(c_change_40_21, bg = c("green", "brown", "purple"), Xnames = c("Plants", "16S", "Soil" ), cutoff = -1)

c_change_40family_21 = varpart(c_change_deep$change21, all_veg_scores_change40, phylum_scores_change40, soil40_change)
# c_stock_10
plot(c_change_40family_21, bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ))

```
```{r}
c_change_40_21_veg = varpart(c_change_deep$change21, veg_scores_change40, canopy_scores_change40)

plot(c_change_40_21_veg)
```



```{r}

c_change_10_21 = varpart(c_change_surface$change21, veg_scores_change10, otu_scores_change10, soil10_change)
# c_stock_10
plot(c_change_10_21, bg = c("green", "brown", "purple"), Xnames = c("Plants", "16S", "Soil Texture" ), cutoff = -1)


c_change_10family_21 = varpart(c_change_surface$change21, veg_scores_change10, phylum_scores_change10, soil10_change)
# c_stock_10
plot(c_change_10family_21, bg = c("green", "brown", "purple"), Xnames = c("Veg", "OTU", "soil" ))

```

```{r}
# otu_data10 = bind_cols(c_10, otuNMDS_scores10)
# otu_model_10 = lm(c_2018 ~ NMDS1 + NMDS2, data = otu_data10)
# summary(otu_model_10)
# 
# 
# 
# otu_data40 = bind_cols(c_40, otuNMDS_scores40)
# otu_model_40 = lm(c_2018 ~ NMDS1 + NMDS2, data = otu_data40)
# summary(otu_model_40)

```
Veg only    

```{r}



```

